---
title: "Diversity project actual analyses"
layout: "post"
tags: "diversity"
output: html_document
editor_options: 
  chunk_output_type: console
---

### Cleaning up and pre-processing

```{r setup, include=F}
knitr::opts_knit$set(root.dir="~/OneDrive - St Vincent's Institute/Documents/RNA\ Diversity/", echo=T)
# But we also have to do this manually because R studio is stupid: 
setwd(knitr::opts_knit$get("root.dir"))
library(ggplot2)
library(rols)
library(data.table)
library(plyr)
library(reshape)
library(tidyverse)
library(viridis)

allSRAFinal <- readRDS("allSRAFinal.rds")

```

### Quick look at the tissue type descriptors to get some numbers... Some of these clearly do not contain cell type info, but others do, and sorting those out is joyous. 

```{r tissues-quick-look}
head(sort(table(allSRAFinal$Cell_line), decreasing=T)) # Some of these are named, some are not... worth keeping that in mind....

table(allSRAFinal$Cell_type) # Useful
table(allSRAFinal$Tissue) # Useful
table(allSRAFinal$tissue_type) # Useful
table(allSRAFinal$Sample_type) # Somewhat useful
table(allSRAFinal$Sample.Name) # Nope
table(allSRAFinal$source_name) # Useful
table(allSRAFinal$source_material_cell_type) # goes with the next column
table(allSRAFinal$source_material) #
table(allSRAFinal$SAMPLE_TYPE) # Somewhat useful
table(allSRAFinal$Sampling_site) # A single study once again...

# The most useful ones are gonna be Tissue, Cell_type and source_name, I think
(table(is.na(allSRAFinal$Tissue), is.na(allSRAFinal$Cell_type)))
head(allSRAFinal[!is.na(allSRAFinal$Tissue) & !is.na(allSRAFinal$Cell_type), ])

(table(is.na(allSRAFinal$Tissue), is.na(allSRAFinal$source_name)))
head(allSRAFinal[!is.na(allSRAFinal$Tissue) & !is.na(allSRAFinal$source_name), ])

(table(is.na(allSRAFinal$Cell_type), is.na(allSRAFinal$source_name)))
head(allSRAFinal[!is.na(allSRAFinal$Cell_type) & !is.na(allSRAFinal$source_name), ])
```

The reason I went down this rabbit hole was that there were about 3000 samples that didn't have tissue info if we focused on the most complete categories: `Cell_type` and `Tissue`. I checked some manually and most of them seemed to have info in `source_name` and `Organism_part`, which I then rescued from the original SRA downloads... so let's look at whether we really need the info from the other columns too:

```{r tissues-quick-look-2}
head(allSRAFinal[is.na(allSRAFinal$Tissue) & is.na(allSRAFinal$Cell_type), ])
head(allSRAFinal[is.na(allSRAFinal$Tissue) & is.na(allSRAFinal$Cell_type) & is.na(allSRAFinal$source_name), ])

dim(allSRAFinal[is.na(allSRAFinal$Tissue) & is.na(allSRAFinal$Cell_type) & is.na(allSRAFinal$source_name) & is.na(allSRAFinal$tissue_type) & is.na(allSRAFinal$Organism_part), ])

noInfo <- allSRAFinal[is.na(allSRAFinal$Tissue) & is.na(allSRAFinal$Cell_type) & is.na(allSRAFinal$source_name) & is.na(allSRAFinal$tissue_type) &is.na(allSRAFinal$Organism_part), ]
table(noInfo$SRA.Study) 

by(noInfo, noInfo$SRA.Study, function(x) head(x, n=20)) 

```

The only one there is the LCLs, ERP001942, which we'll fix later. Looking at the table there's a lot of redundancy between columns, or a lot of studies that report different kinds of info in different columns, so it is worth being strategic so we're not trying to disambiguate a bunch of weird things. 

Once again, false/false is where we want to be looking

```{r tissues-quick-look-3}
table(is.na(allSRAFinal$Tissue), is.na(allSRAFinal$Cell_type)) # Fair amount of overlap
table(is.na(allSRAFinal$Tissue), is.na(allSRAFinal$source_name)) # Lots of overlap
table(is.na(allSRAFinal$Tissue), is.na(allSRAFinal$Organism_part)) # No one uses both of these at the same time

table(is.na(allSRAFinal$Cell_type), is.na(allSRAFinal$source_name))
table(is.na(allSRAFinal$Cell_type), is.na(allSRAFinal$Organism_part))
```

`source_name` is clearly the most redundant one, but I grabbed all of them anyhow and did some manual spotchecking and cleaning up of the term list, which I think is mostly ok/ok enough that results won't change too much. We'll want to keep an eye out on the disease columns later on, though, and on the new disease column I made from those...

```{r write-tissue-descriptors, include=F}
write.table(as.data.frame(table(allSRAFinal$Tissue)), file="Tissue_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")
write.table(as.data.frame(table(allSRAFinal$Cell_type)), file="Cell_type_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")
write.table(as.data.frame(table(allSRAFinal$source_name)), file="source_name_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")
write.table(as.data.frame(table(allSRAFinal$tissue_type)), file="tissue_type_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")
write.table(as.data.frame(table(allSRAFinal$Organism_part)), file="Organism_part_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")
```

Now we explore a few ways of merging data... First we sanity check the order in which to merge...

```{r tissues-merge-order, include=F}
# Merge the two tissue type columns:
allSRAFinal$twoTissue <- coalesce(allSRAFinal$Tissue, allSRAFinal$tissue_type) 
length(table(allSRAFinal$twoTissue)) # A few less than when you sum the other two together, but that's expected.
table(is.na(allSRAFinal$twoTissue)) # 18k have some sort of tissue assignment

allSRAFinal$threeTissue <- coalesce(allSRAFinal$twoTissue, allSRAFinal$Organism_part)
length(table(allSRAFinal$threeTissue)) # A few less than when you sum the other two together, but that's expected.
table(is.na(allSRAFinal$threeTissue)) # only gained like 600... 

allSRAFinal$fourTissue <- coalesce(allSRAFinal$threeTissue, allSRAFinal$Cell_type)
length(table(allSRAFinal$fourTissue)) # A few less than when you sum the other two together, but that's expected.
table(is.na(allSRAFinal$fourTissue)) # 21k have some sort of tissue assignment

allSRAFinal$fiveTissue <- coalesce(allSRAFinal$fourTissue, allSRAFinal$source_name)
length(table(allSRAFinal$fiveTissue)) # A few less than when you sum the other two together, but that's expected.
table(is.na(allSRAFinal$fiveTissue)) # 21k have some sort of tissue assignment

# The missing 464 are the LCLS from that one paper, so we're good there, we'll fix them later. 

# What kind of gaps did we fill with each step?
head(allSRAFinal[!is.na(allSRAFinal$twoTissue) & is.na(allSRAFinal$Tissue),], n = 20)
head(allSRAFinal[!is.na(allSRAFinal$threeTissue) & is.na(allSRAFinal$twoTissue),], n = 20)
head(allSRAFinal[!is.na(allSRAFinal$fourTissue) & is.na(allSRAFinal$threeTissue),], n = 20)
head(allSRAFinal[!is.na(allSRAFinal$fiveTissue) & is.na(allSRAFinal$fourTissue),], n = 20)

head(sort(table(allSRAFinal$twoTissue), decreasing = T)) #
head(sort(table(allSRAFinal$threeTissue), decreasing = T)) #
head(sort(table(allSRAFinal$fourTissue), decreasing = T)) #
head(sort(table(allSRAFinal$fiveTissue), decreasing = T)) #
```

# And now it's time to update each of those, just for sanity checking, before we pick which approach to go with

For some reason the match command won't work well with the indexing inside the command, so we gotta separate it out. Don't care to troubleshoot. 

```{r tissues-merge-update}
cleanInfo <- read.csv("20240616_cell_type_tissue_descriptors.csv")

cleanTissue <- cleanInfo[cleanInfo$coded.as %in% "tissue",]
cleanTissue_type <- cleanInfo[cleanInfo$coded.as %in% "tissue_type",]
cleanOrganism_part <- cleanInfo[cleanInfo$coded.as %in% "Organism_part",]
cleanCell_type <- cleanInfo[cleanInfo$coded.as %in% "cell_type",]
cleanSource_name <- cleanInfo[cleanInfo$coded.as %in% "source_name",]

# This is going to be really annoying, but I think it's easiest to do it by coalescing five different updated columns at the end... except it's ten, because we do system and organ.

allSRAFinal$tissueSystem <- cleanTissue[match(allSRAFinal$Tissue, cleanTissue$Term, incomparables = NA, nomatch = NA),]$system
allSRAFinal$tissue_typeSystem <- cleanTissue_type[match(allSRAFinal$tissue_type, cleanTissue_type$Term, incomparables = NA, nomatch = NA),]$system
allSRAFinal$Organism_partSystem <- cleanOrganism_part[match(allSRAFinal$Organism_part, cleanOrganism_part$Term, incomparables = NA, nomatch = NA),]$system
allSRAFinal$Cell_typeSystem <- cleanCell_type[match(allSRAFinal$Cell_type, cleanCell_type$Term, incomparables = NA, nomatch = NA),]$system
allSRAFinal$source_nameSystem <- cleanSource_name[match(allSRAFinal$source_name, cleanSource_name$Term, incomparables = NA, nomatch = NA),]$system

allSRAFinal$tissueOrgan <- cleanTissue[match(allSRAFinal$Tissue, cleanTissue$Term, incomparables = NA, nomatch = NA),]$organ
allSRAFinal$tissue_typeOrgan <- cleanTissue_type[match(allSRAFinal$tissue_type, cleanTissue_type$Term, incomparables = NA, nomatch = NA),]$organ
allSRAFinal$Organism_partOrgan <- cleanOrganism_part[match(allSRAFinal$Organism_part, cleanOrganism_part$Term, incomparables = NA, nomatch = NA),]$organ
allSRAFinal$Cell_typeOrgan <- cleanCell_type[match(allSRAFinal$Cell_type, cleanCell_type$Term, incomparables = NA, nomatch = NA),]$organ
allSRAFinal$source_nameOrgan <- cleanSource_name[match(allSRAFinal$source_name, cleanSource_name$Term, incomparables = NA, nomatch = NA),]$organ

# And now we coalesce in the order we established above...
allSRAFinal$finalSystem <- coalesce(allSRAFinal$tissueSystem, allSRAFinal$tissue_typeSystem, allSRAFinal$Organism_partSystem, allSRAFinal$Cell_typeSystem, allSRAFinal$source_nameSystem)

allSRAFinal$finalOrgan <- coalesce(allSRAFinal$tissueOrgan, allSRAFinal$tissue_typeOrgan, allSRAFinal$Organism_partOrgan, allSRAFinal$Cell_typeOrgan, allSRAFinal$source_nameOrgan)
```

So that seems like it should have worked, but we should sanity check it a bit more... First, do things agree across the descriptor levels, and if not, what should we do?

```{r tissues-merge-check-figures}
plotSystemTissue <- table(allSRAFinal$finalSystem, allSRAFinal$tissueSystem) %>% melt (.) 
ggplot(plotSystemTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("System annotation") +
  xlab("Tissue") +
  ylab("Final") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plotSystemTissue <- table(allSRAFinal$finalSystem, allSRAFinal$tissue_typeSystem) %>% melt (.) 
ggplot(plotSystemTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("System annotation") +
  xlab("tissue_type") +
  ylab("Final") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plotSystemTissue <- table(allSRAFinal$finalSystem, allSRAFinal$Organism_partSystem) %>% melt (.) 
ggplot(plotSystemTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("System annotation") +
  xlab("Organism_part") +
  ylab("Final") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Yeeeeeep these two are fucked up. Some of it will have been my old approach, but other stuff is gonna be harder to clean up. 
plotSystemTissue <- table(allSRAFinal$finalSystem, allSRAFinal$Cell_typeSystem) %>% melt (.) 
ggplot(plotSystemTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("System annotation") +
  xlab("Cell_type") +
  ylab("Final") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plotSystemTissue <- table(allSRAFinal$finalSystem, allSRAFinal$source_nameSystem) %>% melt (.) 
ggplot(plotSystemTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("System annotation") +
  xlab("source_name") +
  ylab("Final") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# And now same for organs:
plotOrganTissue <- table(allSRAFinal$finalOrgan, allSRAFinal$tissueOrgan) %>% melt (.) 
ggplot(plotOrganTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("Organ annotation") +
  xlab("Tissue") +
  ylab("Final") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plotOrganTissue <- table(allSRAFinal$finalOrgan, allSRAFinal$tissue_typeOrgan) %>% melt (.) 
ggplot(plotOrganTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("Organ annotation") +
  xlab("tissue_type") +
  ylab("Final") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plotOrganTissue <- table(allSRAFinal$finalOrgan, allSRAFinal$Organism_partOrgan) %>% melt (.) 
ggplot(plotOrganTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("Organ annotation") +
  xlab("Organism_part") +
  ylab("Final") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Yeeeeeep these two are fucked up. Some of it will have been my old approach, but other stuff is gonna be harder to clean up. 
plotOrganTissue <- table(allSRAFinal$finalOrgan, allSRAFinal$Cell_typeOrgan) %>% melt (.) 
ggplot(plotOrganTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("Organ annotation") +
  xlab("Cell_type") +
  ylab("Final") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plotOrganTissue <- table(allSRAFinal$finalOrgan, allSRAFinal$source_nameOrgan) %>% melt (.) 
ggplot(plotOrganTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("Organ annotation") +
  xlab("source_name") +
  ylab("Final") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# How often do these two agree?
plotOrganTissue <- table(allSRAFinal$Cell_typeOrgan, allSRAFinal$source_nameOrgan) %>% melt (.) 
ggplot(plotOrganTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("Organ annotation") +
  xlab("source_name") +
  ylab("Cell_type") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# How often do these two agree?
plotOrganSystem <- table(allSRAFinal$finalOrgan, allSRAFinal$finalSystem) %>% melt (.) 
ggplot(plotOrganTissue, aes(Var.2, Var.1)) +
  geom_tile(aes(fill = log2(value)), colour = "white") +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("Master annotation") +
  xlab("System") +
  ylab("Organ") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Uh oh... There's a fair bit of disagreement when we compare the cell type based assignments and the tissue based assignments. And some kind of make sense, and some really don't, like, intestine/iPSC or reproductive/digestive

```{r tissues-merge-update-2}
# Let's look only at the discrepancies... write a little function to show us because otherwise it gets annoying
seeMismatches <- function(Var1, Var2){
  messySubset <- allSRAFinal[allSRAFinal[[Var1]] != allSRAFinal[[Var2]],] %>% .[!is.na(.[[Var2]]),]
  return(messySubset)
  }

cellOrganMismatches <- seeMismatches("finalOrgan", "Cell_typeOrgan")
cellSystemMismatches <- seeMismatches("finalSystem", "Cell_typeSystem")
sourceOrganMismatches <- seeMismatches("finalOrgan", "source_nameOrgan")
sourceSystemMismatches <- seeMismatches("finalSystem", "source_nameSystem")

table(cellOrganMismatches$finalOrgan, cellOrganMismatches$Cell_typeOrgan)
table(cellSystemMismatches$finalSystem, cellSystemMismatches$Cell_typeSystem)
table(sourceOrganMismatches$finalOrgan, sourceOrganMismatches$source_nameOrgan)
table(sourceSystemMismatches$finalSystem, sourceSystemMismatches$source_nameSystem)

# Make my life easier
columnsILike <- c(1:3, 6,7,8,13,26,29,44,45,49,50,51,52)
 
# by(cellOrganMismatches, c(cellOrganMismatches$finalOrgan), function(x) head(x[,columnsILike], n = 20))
# by(cellSystemMismatches, c(cellSystemMismatches$Cell_typeSystem), function(x) head(x[,columnsILike], n = 20))
# by(sourceOrganMismatches, c(sourceOrganMismatches$finalOrgan), function(x) head(x[,columnsILike], n = 20))
# by(sourceSystemMismatches, c(sourceSystemMismatches$finalSystem), function(x) head(x[,columnsILike], n = 20))

table(cellOrganMismatches$SRA.Study)
table(cellSystemMismatches$SRA.Study)
table(sourceOrganMismatches$SRA.Study)
table(sourceSystemMismatches$SRA.Study)

by(cellOrganMismatches, c(cellOrganMismatches$SRA.Study), function(x) head(x[,columnsILike], n = 20))
by(cellSystemMismatches, c(cellSystemMismatches$SRA.Study), function(x) head(x[,columnsILike], n = 20))
by(sourceOrganMismatches, c(sourceOrganMismatches$SRA.Study), function(x) head(x[,columnsILike], n = 20))
by(sourceSystemMismatches, c(sourceSystemMismatches$SRA.Study), function(x) head(x[,columnsILike], n = 20))
```

Some of these are clear incorrect assignments when cleaning up the tissue stuff, and so the big spreadsheet needs to be updated, eg, the foreskin fibroblasts from Coriell which came up as urogenital (SRP070663, now fixed). However, there's also a lot of problems with samples that do end up in different classes depending of which column one looks at, eg, SRP266877 should really be under iPSC but isn't because tissue is given as blood, even though it's an established cell line... Not all the cases are clear cut, and some are straight up contradictory, eg, DRP001797, which says endocardial cells but also endothelium of blood vessels, or, SRP056287, which gives cell type as cortical neurons, and tissue as fibroblast and which seem to be iPSC-derived cortical neurons (updated), or SRP259538, which lists cardiomyocyte and foetal lung (but they're cell-line derived anyhow)

There's a lot of immune cells sampled in the lung, and I'm not sure what to do with them - keep them as both things?

So that's a first pass, but we need to get rid of established cell lines and then come back to some of this. Next, some checks for missingness etc

```{r tissues-merge-update-missing}
table(allSRAFinal$finalSystem)
table(allSRAFinal$finalOrgan)

table(is.na(allSRAFinal$finalSystem), is.na(allSRAFinal$finalOrgan))
table(is.na(allSRAFinal$finalSystem)) 
table(is.na(allSRAFinal$finalOrgan))

noAssignment <- allSRAFinal[is.na(allSRAFinal$finalOrgan) & is.na(allSRAFinal$finalSystem),]
table(noAssignment$SRA.Study)
by(noAssignment, noAssignment$SRA.Study, function(x) head(x))
```

Most of these had info in one of the tissue columns, but sometimes I had to go back to run selector and dig a bit deeper. 

* DRP001150: cell mixture - tissue sample (yikes): no fix, will get dropped
* ERP001942: LCLs
* ERP122083: more cancer: finalOrgan:cancer and finalTissue:cancer
* ERP122103: single cancer sample: finalOrgan:cancer and finalTissue:cancer
* ERP122256: soft tissue cancer: finalOrgan:cancer and finalTissue:cancer
* SRP151215: LCLs
* SRP154973: 'adjacent normal' to cancers, lung cell carcinoma: finalOrgan:lung, finalSystem:respiratory
* SRP179998: brain sample that for some reason isn't matching well: finalOrgan:brain and finalSystem:nervous
* SRP181079: BCR sequencing of B-cells; finalOrgan:blood and finalSystem:immune
* SRP181649: single established cell line: no fix, will get dropped
* SRP186687: established cancer cell lines: no fix, will get dropped
* SRP343308: Intestinal biopsies from multiple tissues: finalOrgan:intestine and finalSystem:digestive
* SRP344328: Intrahepatic cholangiocarcinoma: finalOrgan: liver and finalSystem:hepatic

```{r tissues-merge-update-missing-2}
# What are the NAs...? 464 Of them are the LCL study, so let's fix that, first of all:
allSRAFinal[allSRAFinal$SRA.Study %in% c("ERP001942", "SRP151215", "SRP181079"),]$finalSystem <- "immune"
allSRAFinal[allSRAFinal$SRA.Study %in% c("ERP001942", "SRP151215", "SRP181079"),]$finalOrgan <- "blood"

allSRAFinal[allSRAFinal$SRA.Study %in% c("ERP122083", "ERP122103", "ERP122256"),]$finalSystem <- "cancer"
allSRAFinal[allSRAFinal$SRA.Study %in% c("ERP122083", "ERP122103", "ERP122256"),]$finalOrgan <- "cancer"

allSRAFinal[allSRAFinal$SRA.Study %in% "SRP344328",]$finalSystem <- "hepatic"
allSRAFinal[allSRAFinal$SRA.Study %in% "SRP344328",]$finalOrgan <- "liver"

allSRAFinal[allSRAFinal$SRA.Study %in% "SRP343308",]$finalSystem <- "digestive"
allSRAFinal[allSRAFinal$SRA.Study %in% "SRP343308",]$finalOrgan <- "intestine"

allSRAFinal[allSRAFinal$SRA.Study %in% "SRP179998",]$finalSystem <- "nervous"
allSRAFinal[allSRAFinal$SRA.Study %in% "SRP179998",]$finalOrgan <- "brain"

allSRAFinal[allSRAFinal$SRA.Study %in% "SRP154973",]$finalSystem <- "respiratory"
allSRAFinal[allSRAFinal$SRA.Study %in% "SRP154973",]$finalOrgan <- "lung"

noAssignment <- allSRAFinal[is.na(allSRAFinal$finalOrgan) & is.na(allSRAFinal$finalSystem),]
table(noAssignment$SRA.Study)
by(noAssignment, noAssignment$SRA.Study, function(x) head(x))

rm(noAssignment)
```

Much better... so we've resolved the NAs, but not the tissue ambiguities, which I think might just have to stay as is. 

Now we look at the rows with info encoded in the `Cell_line` column, in addition to our organ and tissue assignments. Lots of these are cell lines that have ancestry info, but are established/immortalised, which was not included in our initial search... but should it have been? A key culprit is SRP186687, which is almost 1000 cancer samples with sequencing. 

The brutal approach is to simply remove the entire study, and then come back again. But it would be smarter to also whitelist some terms and not just remove the column, or remove any rows with info in the column. SRP347253 has some epic cell line names, for instance, and all the occurrences of LCL etc should be kept - but first gotta check they're not already covered in other columns!

```{r tissues-merge-update-cell-line}
table(!is.na(allSRAFinal$Cell_line))

allCellLine <- allSRAFinal[!is.na(allSRAFinal$Cell_line),]
sort(table(allCellLine$SRA.Study), decreasing=T)

head(sort(table(allCellLine$Cell_line), decreasing=T), n =50) # Some of these are clear keepers, some of the are clear removal. But the question is which is which, and how do I make my life easier? Let's check the big ones first:

by(allCellLine, allCellLine$SRA.Study, function(x) sort(table(x$finalOrgan),  decreasing=T))
by(allCellLine, allCellLine$SRA.Study, function(x) sort(table(x$finalSystem),  decreasing=T))
by(allCellLine, allCellLine$SRA.Study, function(x) sort(table(x$Cell_line),  decreasing=T))


by(allCellLine, allCellLine$SRA.Study, function(x) head(x, n=20))
```

* SRP186687: the huge cancer study, currently assigned to various organs and lots of NA. Unsure
* SRP347253: the big iPSC panel, correctly assigned to iPSCs. ok
* ERP114122: LCLs, (from 1000 Genomes). ok
* SRP344545: LCLs, (from 1000 Genomes). ok
* SRP132693: iPSC, iPSC derived cardiomyocytes and primary heart tissue. ok. 
* SRP154009: iPSC and iPSC derived cardiomyocytes. ok
* SRP323748: looks like skin and blood (starting iPSC material), SRA accession says they're all iPSC derived cardiomyocytes from iPSC core, and this is recorded in the 'cell_subtype' column. Need to update tissue. Unsure.
* SRP362734: colorectal cancer organoids... currently processed correctly as if colorectal
* SRP106527: CEU LCL sequencing. ok
SRP217300: MIA PaCa, established breast cancer cell line. Unsure  
SRP090531
SRP125571
SRP151215
SRP103111
SRP174206
SRP223674
SRP330940: skin fibroblasts with known name. Unsure
SRP221374 
SRP075592
SRP237644
SRP408226: HepG2. exclude.
SRP292295
SRP191042
SRP222944
SRP292536
SRP056109
SRP073347 
SRP367525: primary vaginal samples. ok.
SRP266877
SRP269147
SRP155367
SRP178235
SRP179498
SRP193979
SRP259538
SRP334792: weird names, need to check. 
SRP352928: LCLs. ok.
SRP131505
SRP198410
SRP219483
SRP220383
SRP091668
SRP041840
SRP057446
SRP060253 
SRP194087
SRP220267: Caco2. exclude. 
SRP070663
SRP052896
SRP066625
SRP181649
SRP298738: established AML cell line. exclude
SRP074707
SRP111915 
SRP164930 

studies to exclude:

c("SRP298738", "SRP220267", "SRP408226")





```{r tissues-merge-update-3}

# Make my life easier
columnsILike <- c(1:3, 6,7,8,13,26,29,51, 52)

# And a bit of a visual sanity check...
by(allSRAFinal, allSRAFinal$finalSystem, function(x) head(x[,columnsILike], n=10))
by(allSRAFinal, allSRAFinal$finalOrgan, function(x) head(x[,columnsILike], n=10))
```


The umbilical cord is incorrectly classified as reproductive, when it's MSCs, so let's be systematic and ask about disagreement between the all possible classifications:



Stopping here for tonight... Why do some classifications disagree sometimes? Some of these are expected, but we should decide which classification we're using, I guess - system or organ, but also, cell type or tissue?? 

```{r check-disagreements}

table(allSRAFinal$system == allSRAFinal$system.CL)
table(allSRAFinal$organ == allSRAFinal$organ.CL)
table(allSRAFinal$refined == allSRAFinal$refined.CL)

head(allSRAFinal$refined == allSRAFinal$refined.CL)



```


There are indeed some massive lung data sets which are probably single cell (similar to the LN dataset I noticed while processing), but anyhow, looks promising. Now time to get looking at the data along both dimensions at the same time:

First, some summary statistics

```{r ancestry-tissue}
# And now that we have cleaned that up... let's do something glorious with it:

# First we focus on population descriptors:
bigAncestrySummary <- allSRAFinal %>% count(SRA.Study, jointAncestry, UBERON, system, organ)
bigRaceSummary <- allSRAFinal %>% count(SRA.Study, jointRace, UBERON, system, organ)
ancestrySummary <- allSRAFinal %>% count(jointAncestry, UBERON, system, organ)
raceSummary <- allSRAFinal %>% count(jointRace, UBERON, system, organ)

meltAncestry <- melt(ancestrySummary)
meltRace <- melt(raceSummary)

```
And now a lot of plots:

```{r plots-plots-plots, echo=F}
ggplot(meltAncestry, aes(x = system, y=value, fill=jointAncestry)) +
  geom_bar(stat="identity") +
  ggtitle("Biological system by ancestry") +
  # xlab("System") +
  ylab("Count") +
  theme_bw() +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(meltAncestry, aes(x = organ, y= value, fill=jointAncestry)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Organ by ancestry") +
  # xlab("Organ") +
  ylab("Count") +
  scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(meltAncestry, aes(x = UBERON, y= value, fill=jointAncestry)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("UBERON term by ancestry") +
  # xlab("UBERON") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(meltAncestry, aes(fill = system, y= value, x=jointAncestry)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Biological system by ancestry") +
  # xlab("System") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(meltAncestry, aes(fill = organ, y= value, x=jointAncestry)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Organ by ancestry") +
  # xlab("Organ") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(meltAncestry, aes(fill = UBERON, y= value, x=jointAncestry)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("UBERON term by ancestry") +
  # xlab("UBERON") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(meltRace, aes(x = system, y= value, fill=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Biological system by Race") +
  # xlab("System") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(meltRace, aes(x = organ, y= value, fill=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Organ by Race") +
  # xlab("System") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(meltRace, aes(x = UBERON, y= value, fill=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Organ by Race") +
  # xlab("UBERON") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(meltRace, aes(fill = system, y= value, x=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Biological system by Race") +
  # xlab("System") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(meltRace, aes(fill = organ, y= value, x=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Organ by Race") +
  # xlab("Organ") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(meltRace, aes(fill = UBERON, y= value, x=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("UBERON term by Race") +
  # xlab("UBERON") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

The same plots, but without NAs... 

```{r plots-plots-plots-na, echo=F}
meltAncestry %>% drop_na(c(system, jointAncestry)) %>%
ggplot(., aes(x = system, y=value, fill=jointAncestry)) +
  geom_bar(stat="identity") +
  ggtitle("Biological system by ancestry") +
  # xlab("System") +
  ylab("Count") +
  theme_bw() +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

meltAncestry %>% drop_na(c(organ, jointAncestry)) %>%
ggplot(., aes(x = organ, y= value, fill=jointAncestry)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Organ by ancestry") +
  # xlab("Organ") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

meltAncestry %>% drop_na(c(UBERON, jointAncestry)) %>%
ggplot(., aes(x = UBERON, y= value, fill=jointAncestry)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("UBERON term by ancestry") +
  # xlab("UBERON") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

meltAncestry %>% drop_na(c(system, jointAncestry)) %>%
ggplot(., aes(fill = system, y= value, x=jointAncestry)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Biological system by ancestry") +
  # xlab("System") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

meltAncestry %>% drop_na(c(organ, jointAncestry)) %>%
ggplot(., aes(fill = organ, y= value, x=jointAncestry)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Organ by ancestry") +
  # xlab("Organ") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

meltAncestry %>% drop_na(c(UBERON, jointAncestry)) %>%
ggplot(., aes(fill = UBERON, y= value, x=jointAncestry)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("UBERON term by ancestry") +
  # xlab("UBERON") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

meltRace %>% drop_na(c(system, jointRace)) %>%
ggplot(., aes(x = system, y= value, fill=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Biological system by Race") +
  # xlab("System") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

meltRace %>% drop_na(c(organ, jointRace)) %>%
ggplot(., aes(x = organ, y= value, fill=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Organ by Race") +
  # xlab("System") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

meltRace %>% drop_na(c(UBERON, jointRace)) %>%
ggplot(., aes(x = UBERON, y= value, fill=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Organ by Race") +
  # xlab("UBERON") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

meltRace %>% drop_na(c(system, jointRace)) %>%
ggplot(., aes(fill = system, y= value, x=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Biological system by Race") +
  # xlab("System") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

meltRace %>% drop_na(c(organ, jointRace)) %>%
ggplot(., aes(fill = organ, y= value, x=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Organ by Race") +
  # xlab("Organ") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

meltRace %>% drop_na(c(UBERON, jointRace)) %>%
ggplot(., aes(fill = UBERON, y= value, x=jointRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("UBERON term by Race") +
  # xlab("UBERON") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We're also interested in how diverse a given study is, and how many studies include samples from each ancestry. We can easily calculate all of that too, although it is hard to see visually:

```{r ancestry-by-study}
ancestryStudy <- allSRAFinal %>% count(SRA.Study, jointAncestry) %>% drop_na(jointAncestry)

# How many studies with any sort of ancestry info?
dim(ancestryStudy)
length(unique(ancestryStudy$SRA.Study))
# How many studies contain samples from each ancestry?
table(ancestryStudy$jointAncestry) 

# And some quick stats on diversity by study:
ancestryStudy %>% count(SRA.Study) %>% summary()

# But... how many samples?
ggplot(ancestryStudy, aes(x = jointAncestry, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  ggtitle("Ancestry by Study") +
  ylab("Count") +
  scale_fill_viridis(discrete=T, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

# This is hard to see, so:
ddply(ancestryStudy, "jointAncestry", summarise, totalN = sum(n), meanN=mean(n), maxN=max(n), nStudies=length(unique((SRA.Study))))

# Now adding the organ system dimension, can repeat with organ and uberon but for now it's ok
ancestryStudySystem <- allSRAFinal %>% count(SRA.Study, jointAncestry, system) %>% drop_na(jointAncestry, system)
dim(ancestryStudySystem)
length(unique(ancestryStudySystem$SRA.Study))

ggplot(ancestryStudySystem, aes(x = jointAncestry, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  facet_wrap( ~ system, nrow=3) +
  ggtitle("Ancestry by study by organ system") +
  # xlab("UBERON") +
  ylab("Count") +
  scale_fill_viridis(discrete=T, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

ggplot(ancestryStudySystem, aes(x = system, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  facet_wrap( ~ jointAncestry, nrow=3) +
  ggtitle("Ancestry by study by organ system") +
  # xlab("UBERON") +
  ylab("Count") +
  scale_fill_viridis(discrete=T, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

# And now really broken down:
ddply(ancestryStudySystem, c("jointAncestry", "system"), summarise, totalN = sum(n), meanN=mean(n), maxN=max(n), nStudies=length(unique((SRA.Study))))

# There's a lot of Asian and East Asian blood! Where is it coming from?
ancestryStudySystem[ancestryStudySystem$system %in% "blood",] %>% arrange(n) %>% tail(n=10)
```

* SRP245400: Splicing profile by capture RNA-seq identifies pathogenic germline variants in tumor suppressor genes (Ambry Genetics)
* SRP333281: Intergrative Multi-Omics of Asian Pregnant Mother (Singapore)
* SRP333281: Intergrative Multi-Omics of Asian Pregnant Mother (Singapore)
* SRP237697: mRNA sequencing of 28 visceral leishmaniasis-HIV patients in NW-Ethiopia (pre-post treatment)
* SRP388678: Identification and interpretation of splicing variants using paired DNA and RNA sequencing in a clinical cohort of patients tested for hereditary cancer predisposition (Ambry Genetics)
* SRP358657: Local and systemic transcriptomic responses from acute exercise induced muscle damage of the human knee extensor
* SRP126583: A modular transcriptional signature identifies phenotypic heterogeneity of human tuberculosis infection
* SRP126691: A modular transcriptional signature identifies phenotypic heterogeneity of human tuberculosis infection
* SRP333281: Intergrative Multi-Omics of Asian Pregnant Mother (Singapore)
* SRP200298: Transcriptome and methylome of human suicide attempter and major depressive disorder patient (Korea)

These make sense! The Singaporean cohorts will be the three main ancestry: (South) Chinese, Malay and Tamil; Ambry is of course invested in diversity, TB is unlikely to show up in Europe. Also I am willing to bet any amount of money that the digestive Asian sequencing comes from cancer samples too somehow?

Anyhow, now we do the same for race:

```{r race-by-study}
RaceStudy <- allSRAFinal %>% count(SRA.Study, jointRace) %>% drop_na(jointRace)

# How many studies with any sort of Race info?
dim(RaceStudy)
length(unique(RaceStudy$SRA.Study))
# How many studies contain samples from each Race?
table(RaceStudy$jointRace) 

# And some quick stats on diversity by study:
RaceStudy %>% count(SRA.Study) %>% summary()

# But... how many samples?
ggplot(RaceStudy, aes(x = jointRace, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  ggtitle("Race by Study") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

# This is hard to see, so:
ddply(RaceStudy, "jointRace", summarise, totalN = sum(n), meanN=mean(n), maxN=max(n), nStudies=length(unique((SRA.Study))))

# Now adding the organ system dimension, can repeat with organ and uberon but for now it's ok
RaceStudySystem <- allSRAFinal %>% count(SRA.Study, jointRace, system) %>% drop_na(jointRace, system)
dim(RaceStudySystem)
length(unique(RaceStudySystem$SRA.Study))

ggplot(RaceStudySystem, aes(x = jointRace, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  facet_wrap( ~ system, nrow=3) +
  ggtitle("Race by study by organ system") +
  # xlab("UBERON") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

ggplot(RaceStudySystem, aes(x = system, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  facet_wrap( ~ jointRace, nrow=3) +
  ggtitle("Race by study by organ system") +
  # xlab("UBERON") +
  ylab("Count") +
  scale_fill_viridis(discrete=T,, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

# And now really broken down:
ddply(RaceStudySystem, c("jointRace", "system"), summarise, totalN = sum(n), meanN=mean(n), maxN=max(n), nStudies=length(unique((SRA.Study))))
```
Wowowowow the stark difference. Would be nice to slice this by country of sampling to see if this is really driven by the USA, or if random people are using race terms.
