---
title: "Diversity project analyses first pass"
layout: "post"
tags: "diversity"
output: html_document
editor_options: 
  chunk_output_type: console
---

Omg finally ready to analyse data! There's still some things to clean up, but I'm too tired to do them...

We'll have to clean up the studies that have info on both race and geography because of shoddy misclassifications...

```{r setup, include=F}
knitr::opts_knit$set(root.dir="~/OneDrive - St Vincent's Institute/Documents/RNA\ Diversity/", echo=T)
# But we also have to do this manually because R studio is stupid: 
setwd(knitr::opts_knit$get("root.dir"))
library(ggplot2)
library(data.table)
library(plyr)
library(reshape)
library(tidyverse)
library(viridis)

allSRAFinal <- readRDS("allSRAFinalTissues.rds")
```
First, some summary statistics

```{r Geography-tissue}
# And now that we have cleaned that up... let's do something glorious with it:

# First we focus on population descriptors:
bigGeographySummary <- allSRAFinal %>% count(SRA.Study, finalGeography, finalSystem, finalOrgan)
bigRaceSummary <- allSRAFinal %>% count(SRA.Study, finalRace, finalSystem, finalOrgan)
GeographySummary <- allSRAFinal %>% count(finalGeography, finalSystem, finalOrgan)
raceSummary <- allSRAFinal %>% count(finalRace, finalSystem, finalOrgan)

meltGeography <- melt(GeographySummary)
meltRace <- melt(raceSummary)

```

And now a lot of plots, removing the NAs from each respective category... 

```{r plots-plots-plots-na, echo=F}
meltGeography %>% drop_na(c(finalSystem, finalGeography)) %>%
  ggplot(., aes(x = finalSystem, y=value, fill=finalGeography)) +
    geom_bar(stat="identity") +
    ggtitle("Biological System by Geography") +
    # xlab("finalSystem") +
    ylab("Count") +
    theme_bw() +
    scale_fill_viridis(discrete=T, na.value="grey50") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme(legend.position="bottom")

meltGeography %>% drop_na(c(finalOrgan, finalGeography)) %>%
  ggplot(., aes(x = finalOrgan, y= value, fill=finalGeography)) +
    geom_bar(stat="identity") +
    theme_bw() +
    ggtitle("Organ by Geography") +
    # xlab("finalOrgan") +
    ylab("Count") +
    scale_fill_viridis(discrete=T, na.value="grey50") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme(legend.position="bottom")

meltRace %>% drop_na(c(finalSystem, finalRace)) %>%
  ggplot(., aes(x = finalSystem, y= value, fill=finalRace)) +
    geom_bar(stat="identity") +
    theme_bw() +
    ggtitle("Biological finalSystem by Race") +
    # xlab("finalSystem") +
    ylab("Count") +
    scale_fill_viridis(discrete=T, na.value="grey50") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme(legend.position="bottom")

meltRace %>% drop_na(c(finalOrgan, finalRace)) %>%
  ggplot(., aes(x = finalOrgan, y= value, fill=finalRace)) +
    geom_bar(stat="identity") +
    theme_bw() +
    ggtitle("finalOrgan by Race") +
    # xlab("finalSystem") +
    ylab("Count") +
    scale_fill_viridis(discrete=T, na.value="grey50") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme(legend.position="bottom")
```

We're also interested in how diverse a given study is, and how many studies include samples from each Geography. We can easily calculate all of that too, although it is hard to see visually. But first we want to check how many studies ended up mixed between race and geography and do some cleaning:

```{r checking-population-assignments}
tabulatePops <- by(allSRAFinal, allSRAFinal$SRA.Study, function(x) table(x$finalGeography, x$finalRace)) 
conflictStudies <- names(tabulatePops[lapply(tabulatePops, function(x) grep("NULL", dimnames(x))) %>% grepl(0, .)])
length(conflictStudies) # That's a lot of messiness...

by(allSRAFinal[allSRAFinal$SRA.Study %in% conflictStudies,], allSRAFinal[allSRAFinal$SRA.Study %in% conflictStudies,]$SRA.Study, function(x) table(x$finalGeography, x$finalRace))

tabulateTerms <- by(allSRAFinal[allSRAFinal$SRA.Study %in% conflictStudies,], allSRAFinal[allSRAFinal$SRA.Study %in% conflictStudies,]$SRA.Study, function(x) table(is.na(x$finalGeography), is.na(x$finalRace))) # So there's often a clear skew towards one category, which should make solving these easier...

tabulateTerms <- data.frame(melt(unlist(tabulateTerms)))
tabulateTerms$condition <- rep(c("both", "race", "geography", "neither"), nrow(tabulateTerms)/4)
tabulateTerms$SRA.Study <- str_sub(rownames(tabulateTerms), end=-2)

ggplot(tabulateTerms, aes(x = SRA.Study, y = value, fill=condition)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Race or Ethnicity usage") +
  # xlab("finalSystem") +
  ylab("Count") +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position="bottom")
```

From this it's very clear what we should do for most studies! The vast majority of them should use only racial terms and (I'm guessing), swap Asian over to a racial descriptor. But anyhow, let's manually spot check some of these:

```{r checking-population-assignments-2}
columnsILike <- c(1,10, 15, 34:36) # Just need those two intermediate ones tbh

by(allSRAFinal[allSRAFinal$SRA.Study %in% conflictStudies,], allSRAFinal[allSRAFinal$SRA.Study %in% conflictStudies,]$SRA.Study, function(x) x[,columnsILike]) # Yeah... these are generally easy to resolve. Will send an email about them soon. 

```


```{r more-faffing-with-descriptors}
allSRAFinal[allSRAFinal$SRA.Study %in% "ERP116722", columnsILike] # Need to fix "other"
allSRAFinal[allSRAFinal$SRA.Study %in% "ERP117085", columnsILike] # Would defs swap to Subsaharan Africa, study is from UCL
allSRAFinal[allSRAFinal$SRA.Study %in% "SRP226691", columnsILike] # Swap to europe
allSRAFinal[allSRAFinal$SRA.Study %in% "SRP102952", columnsILike] # Swap to white
allSRAFinal[allSRAFinal$SRA.Study %in% "SRP114762", columnsILike] # Should probably be race.
allSRAFinal[allSRAFinal$SRA.Study %in% "SRP116913", columnsILike] # US Study, should be race.
  table(allSRAFinal[allSRAFinal$SRA.Study %in% "SRP116913",]$ETHNICITY)

allSRAFinal[allSRAFinal$SRA.Study %in% "SRP388678", columnsILike] # Not sure what to do with this one... clearly US terms, but also terms outside the usual categories.
```



```{r Geography-by-study, include=F, eval=F}

by(allSRAFinal[allSRAFinal$SRA.Study %in% conflictStudies,], allSRAFinal[allSRAFinal$SRA.Study %in% conflictStudies,]$SRA.Study, function(x) head(x[, columnsILike], n=30))

GeographyStudy <- allSRAFinal %>% count(SRA.Study, finalGeography) %>% drop_na(finalGeography)

# How many studies with any sort of Geography info?
dim(GeographyStudy)
length(unique(GeographyStudy$SRA.Study))
# How many studies contain samples from each Geography?
table(GeographyStudy$finalGeography) 

# And some quick stats on diversity by study:
GeographyStudy %>% count(SRA.Study) %>% summary()

# But... how many samples?
ggplot(GeographyStudy, aes(x = finalGeography, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  ggtitle("Geography by Study") +
  ylab("Count") +
  scale_fill_viridis(discrete=T, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

# This is hard to see, so:
ddply(GeographyStudy, "finalGeography", summarise, totalN = sum(n), meanN=mean(n), maxN=max(n), nStudies=length(unique((SRA.Study))))

# Now adding the finalOrgan finalSystem dimension, can repeat with finalOrgan and uberon but for now it's ok
GeographyStudyfinalSystem <- allSRAFinal %>% count(SRA.Study, finalGeography, finalSystem) %>% drop_na(finalGeography, finalSystem)
dim(GeographyStudyfinalSystem)
length(unique(GeographyStudyfinalSystem$SRA.Study))

ggplot(GeographyStudyfinalSystem, aes(x = finalGeography, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  facet_wrap( ~ finalSystem, nrow=3) +
  ggtitle("Geography by study by finalOrgan finalSystem") +
  # xlab("UBERON") +
  ylab("Count") +
  scale_fill_viridis(discrete=T, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

ggplot(GeographyStudyfinalSystem, aes(x = finalSystem, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  facet_wrap( ~ finalGeography, nrow=3) +
  ggtitle("Geography by study by finalOrgan finalSystem") +
  # xlab("UBERON") +
  ylab("Count") +
  scale_fill_viridis(discrete=T, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

# And now really broken down:
ddply(GeographyStudyfinalSystem, c("finalGeography", "finalSystem"), summarise, totalN = sum(n), meanN=mean(n), maxN=max(n), nStudies=length(unique((SRA.Study))))

# There's a lot of Asian and East Asian blood! Where is it coming from?
GeographyStudyfinalSystem[GeographyStudyfinalSystem$finalSystem %in% "blood",] %>% arrange(n) %>% tail(n=10)
```

* SRP245400: Splicing profile by capture RNA-seq identifies pathogenic germline variants in tumor suppressor genes (Ambry Genetics)
* SRP333281: Intergrative Multi-Omics of Asian Pregnant Mother (Singapore)
* SRP333281: Intergrative Multi-Omics of Asian Pregnant Mother (Singapore)
* SRP237697: mRNA sequencing of 28 visceral leishmaniasis-HIV patients in NW-Ethiopia (pre-post treatment)
* SRP388678: Identification and interpretation of splicing variants using paired DNA and RNA sequencing in a clinical cohort of patients tested for hereditary cancer predisposition (Ambry Genetics)
* SRP358657: Local and finalSystemic transcriptomic responses from acute exercise induced muscle damage of the human knee extensor
* SRP126583: A modular transcriptional signature identifies phenotypic heterogeneity of human tuberculosis infection
* SRP126691: A modular transcriptional signature identifies phenotypic heterogeneity of human tuberculosis infection
* SRP333281: Intergrative Multi-Omics of Asian Pregnant Mother (Singapore)
* SRP200298: Transcriptome and methylome of human suicide attempter and major depressive disorder patient (Korea)

These make sense! The Singaporean cohorts will be the three main Geography: (South) Chinese, Malay and Tamil; Ambry is of course invested in diversity, TB is unlikely to show up in Europe. Also I am willing to bet any amount of money that the digestive Asian sequencing comes from cancer samples too somehow?

Anyhow, now we do the same for race:

```{r race-by-study, include=F, eval=F}
RaceStudy <- allSRAFinal %>% count(SRA.Study, finalRace) %>% drop_na(finalRace)

# How many studies with any sort of Race info?
dim(RaceStudy)
length(unique(RaceStudy$SRA.Study))
# How many studies contain samples from each Race?
table(RaceStudy$finalRace) 

# And some quick stats on diversity by study:
RaceStudy %>% count(SRA.Study) %>% summary()

# But... how many samples?
ggplot(RaceStudy, aes(x = finalRace, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  ggtitle("Race by Study") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

# This is hard to see, so:
ddply(RaceStudy, "finalRace", summarise, totalN = sum(n), meanN=mean(n), maxN=max(n), nStudies=length(unique((SRA.Study))))

# Now adding the finalOrgan finalSystem dimension, can repeat with finalOrgan and uberon but for now it's ok
RaceStudyfinalSystem <- allSRAFinal %>% count(SRA.Study, finalRace, finalSystem) %>% drop_na(finalRace, finalSystem)
dim(RaceStudyfinalSystem)
length(unique(RaceStudyfinalSystem$SRA.Study))

ggplot(RaceStudyfinalSystem, aes(x = finalRace, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  facet_wrap( ~ finalSystem, nrow=3) +
  ggtitle("Race by study by finalOrgan finalSystem") +
  # xlab("UBERON") +
  ylab("Count") +
scale_fill_viridis(discrete=T, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

ggplot(RaceStudyfinalSystem, aes(x = finalSystem, y=n, fill=SRA.Study)) +
  geom_bar(stat="identity") +
  facet_wrap( ~ finalRace, nrow=3) +
  ggtitle("Race by study by finalOrgan finalSystem") +
  # xlab("UBERON") +
  ylab("Count") +
  scale_fill_viridis(discrete=T,, na.value="grey50") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position = "none") 

# And now really broken down:
ddply(RaceStudyfinalSystem, c("finalRace", "finalSystem"), summarise, totalN = sum(n), meanN=mean(n), maxN=max(n), nStudies=length(unique((SRA.Study))))
```
Wowowowow the stark difference. Would be nice to slice this by country of sampling to see if this is really driven by the USA, or if random people are using race terms.
