---
title: "Diversity project actual analyses I swear"
layout: "post"
tags: "diversity"
output: html_document
editor_options: 
  chunk_output_type: console
---

We will finally analyse this data properly. 
```{r setup, include=F}
knitr::opts_knit$set(root.dir="~/OneDrive - St Vincent's Institute/Documents/RNA\ Diversity/", echo=T)
# But we also have to do this manually because R studio is stupid: 
setwd(knitr::opts_knit$get("root.dir"))
library(ggplot2)
library(data.table)
library(plyr)
library(reshape)
library(tidyverse)
library(patchwork)
library(ggalluvial)
options(device = "quartz")

allSRAFinal <- readRDS("20240704_allSRAFinal.rds")
```

After discussion with Amanda, we'll be using the strictestRace/Geography assignments, see `06_revisiting_population_terms` for more context on what this means.

```{r plotting-prep, out.width="100%"}
# First we clean up the plotting labels:
allSRAFinal <- allSRAFinal %>% 
    mutate(strictestRace = gsub("or ", "or\n", strictestRace)) %>%
    mutate(strictestGeography = gsub("and ", "and\n", strictestGeography)) %>%
    mutate(strictestGeography = gsub("^Asia$", "Asia (NOS)", strictestGeography)) %>%
    as.data.frame()

# Let's sort out some colour schemes... For Race we simply use Dark2, but for Geography we are following Alicia Martin and using a somewhat custom palette that seems to borrow from Set1 and Dark2; I opened a paper in Affinity and grabbed the html values and then added a couple for the new entries (Southeast Asia, Asia NOS):

scale_fill_geography <- function(...){
    ggplot2:::manual_scale(
        'fill', 
        values = setNames(c('#555555', '#04C3C8', '#DB6968', '#60BC55', '#9C8DC3', '#F3D78A', '#4D97CD', '#FBC9C4', '#BCCC45', '#F8984E', '#C69C3A', '#8B96AD' ), unique(allSRAFinal$strictestGeography)), 
        ...
    )
}

scale_color_geography <- function(...){
    ggplot2:::manual_scale(
        'color', 
        values = setNames(c('#555555', '#04C3C8', '#DB6968', '#60BC55', '#9C8DC3', '#F3D78A', '#4D97CD', '#FBC9C4', '#BCCC45', '#F8984E', '#C69C3A', '#8B96AD' ), unique(allSRAFinal$strictestGeography)), 
        ...
    )
}

allSRAFinal$strictestGeography <- factor(allSRAFinal$strictestGeography, levels = c("Subsaharan Africa", "North Africa and\nWestern Asia", "Europe", "South Asia", "Southeast Asia", "East Asia", "Asia (NOS)", "Oceania", "Americas", "Multiple", "Other"))

```

## 1. Who is being sequenced?

```{r plot-survey-level, out.width="100%"}
geographyClean <- allSRAFinal %>% drop_na(strictestGeography) %>% count(strictestGeography)
raceClean <- allSRAFinal %>% drop_na(strictestRace) %>% count(strictestRace)

allGeo <- ggplot(geographyClean, aes(x = strictestGeography, y = n, fill = strictestGeography)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Submitted descriptor (ancestral/geographic origin)") +
  xlab("") +
  ylab("Samples") +
  # scale_y_continuous(trans='log10') +
  scale_fill_geography() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none")

allRace <- ggplot(raceClean, aes(x = strictestRace, y = n, fill=strictestRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Submitted descriptor (US Census race term)") +
  xlab("") +
  ylab("Samples") +
  # scale_y_continuous(trans='log10') +
  scale_fill_brewer(palette="Dark2") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none")
```

A bit more granular now, showing n samples per study:

```{r plot-survey-level-2, out.width="100%"}
sampleGeography <- allSRAFinal %>% count(SRA.Study, strictestGeography) %>% drop_na(strictestGeography)
sampleRace <- allSRAFinal %>% count(SRA.Study, strictestRace) %>% drop_na(strictestRace)
meltSampleGeography <- melt(sampleGeography)
meltSampleRace <- melt(sampleRace)

geoStudy <- ggplot(meltSampleGeography, aes(x = strictestGeography, y=value,fill=strictestGeography)) +
  geom_boxplot(width=0.8, outlier.shape = NA) +
  geom_jitter(size=0.5, width=0.2, color="#333333") +
  theme_bw() +
  ggtitle("Submitted descriptor (ancestral/geographic origin)") +
  xlab("") +
  ylab("Samples per study") +
  # scale_y_continuous(trans='log10') +
  scale_fill_geography() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "none")

raceStudy <- ggplot(meltSampleRace, aes(x = strictestRace, y=value,fill=strictestRace)) +
  geom_boxplot(width=0.8, outlier.shape = NA) +
  geom_jitter(size=0.5, width=0.2, color="#333333") +
  theme_bw() +
  ggtitle("Submitted descriptor (US Census race term)") +
  xlab("") +
  ylab("Samples per study") +
  # scale_y_continuous(trans='log10') +
  scale_fill_brewer(palette="Dark2") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "none")
```

## 2. And by who? Where is the sequencing happening?

```{r plot-clean-country, out.width="100%"}
geographyClean <- allSRAFinal %>% drop_na(strictestGeography)
raceClean <- allSRAFinal %>% drop_na(strictestRace)

geoCountry <- ggplot(geographyClean, aes(x = finalCountry, fill = strictestGeography)) +
  geom_bar(position = position_dodge2(width = 0.9, preserve = "single")) +
  theme_bw() +
  ggtitle("SRA Depositor Country") +
  xlab("") +
  ylab("Samples") +
  # scale_y_continuous(trans='log10') +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

raceCountry <- ggplot(raceClean, aes(x = finalCountry, fill=strictestRace)) +
  geom_bar(position = position_dodge2(width = 0.9, preserve = "single")) +
  theme_bw() +
  ggtitle("SRA Depositor Country") +
  xlab("") +
  ylab("Samples") +
  # scale_y_continuous(trans='log10') +
  scale_fill_brewer(palette="Dark2", name="US Census\nRacial Term") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

# Let's attempt an alluvial plot and call it quits:
geographyAlluvial <- allSRAFinal %>% count(strictestGeography, finalCountry) %>% drop_na(c(strictestGeography, finalCountry)) %>%
ggplot(data = .,
       aes(axis1 = finalCountry, axis2 = strictestGeography, y = n)) +
  scale_x_discrete(limits = c("SRA Depositor\nCountry", "Population\nDescriptor"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestGeography)) +
  scale_fill_geography(name="Population\nDescriptor") +
  geom_stratum(width=1/3) +
  ylab("Samples") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  theme(legend.position = "none")

raceAlluvial <- allSRAFinal %>% count(strictestRace, finalCountry) %>% drop_na(c(strictestRace, finalCountry)) %>%
ggplot(data = .,
       aes(axis1 = finalCountry, axis2 = strictestRace, y = n)) +
  scale_x_discrete(limits = c("SRA Depositor\nCountry", "US Census Racial Term"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestRace)) +
  scale_fill_brewer(palette="Dark2", name="US Census\nRacial Term") +
  geom_stratum(width=1/3) +
  ylab("Samples") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  theme(legend.position = "none")

```
## 3. What tissues are they sequencing?

```{r first-pass-tissue, out.width="100%"}
# First we focus on population descriptors:
bigGeographySummary <- allSRAFinal %>% count(SRA.Study, strictestGeography, finalSystem, finalOrgan, finalDisease)
bigRaceSummary <- allSRAFinal %>% count(SRA.Study, strictestRace, finalSystem, finalOrgan, finalDisease)
geographySummary <- allSRAFinal %>% count(strictestGeography, finalSystem, finalOrgan, finalDisease)
raceSummary <- allSRAFinal %>% count(strictestRace, finalSystem, finalOrgan, finalDisease)

meltGeography <- melt(geographySummary)
meltRace <- melt(raceSummary)

geoSystem <- meltGeography %>% drop_na(c(finalSystem, strictestGeography)) %>%
  ggplot(., aes(x = finalSystem, y = value, fill = strictestGeography)) +
  geom_bar(stat="identity") +
  ggtitle("Reported biological system") +
  xlab("") +
  ylab("Samples") +
  theme_bw() +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

geoOrgan <- meltGeography %>% drop_na(c(finalOrgan, strictestGeography)) %>%
  ggplot(., aes(x = finalOrgan, y = value, fill = strictestGeography)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported organ") +
  xlab("") +
  ylab("Samples") +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "bottom")

geoDisease <- meltGeography %>% drop_na(c(finalDisease, strictestGeography)) %>%
  ggplot(., aes(x = finalDisease, y = value, fill = strictestGeography)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported disease") +
  xlab("") +
  ylab("Samples") +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")


raceSystem <- meltRace %>% drop_na(c(finalSystem, strictestRace)) %>%
  ggplot(., aes(x = finalSystem, y = value, fill = strictestRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported biological system") +
  xlab("") +
  ylab("Samples") +
  scale_fill_brewer(palette="Dark2", name="US Census\nRacial Term") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

raceOrgan <- meltRace %>% drop_na(c(finalOrgan, strictestRace)) %>%
  ggplot(., aes(x = finalOrgan, y = value, fill = strictestRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported organ") +
  xlab("") +
  ylab("Samples") +
  scale_fill_brewer(palette="Dark2", name="US Census\nRacial Term") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "bottom")

raceDisease <- meltRace %>% drop_na(c(finalDisease, strictestRace)) %>%
  ggplot(., aes(x = finalDisease, y = value, fill = strictestRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported disease") +
  xlab("") +
  ylab("Samples") +
  scale_fill_brewer(palette="Dark2", name="US Census\nRacial Term") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

```

```{r arrange-plots, out.width="100%"}

pdf(file="all_geography_plots.pdf", height=9, width=9)
  allGeo + geoCountry + geoStudy + geoSystem + geoOrgan + geoDisease +
    plot_layout(nrow=2)
dev.off()

pdf(file="all_race_plots.pdf", height=9, width=9)
  allRace + raceCountry + raceStudy + raceSystem + raceOrgan + raceDisease +
    plot_layout(nrow=2)
dev.off()

pdf(file="bar_vs_alluvial.pdf", height=9, width=9)
  geoCountry + geographyAlluvial + raceCountry + raceAlluvial +
    plot_layout(nrow=2)
dev.off()
```

Well that's nice!

Now let's try to tease out the relationship between things a bit better, even if the knit file is really ugly

## 4. Attempting alluvial plots...

For these (and only for these!), we're gonna set `cancer` in either `finalOrgan` or `finalSystem` to NA, because otherwise the data gets messy. Worth discussing what to do in general/above. 

```{r attempt-alluvial-1, out.width="100%"}
geoTissueAlluvial <- geographySummary %>% filter(!grepl('cancer', finalSystem)) %>% filter(!grepl('cancer', finalOrgan)) %>% drop_na(c(strictestGeography, finalSystem, finalOrgan, finalDisease)) %>%
ggplot(data = .,
       aes(axis1 = finalSystem, axis2 = finalOrgan, axis3 = finalDisease, y = n)) +
  scale_x_discrete(limits = c("Biological system", "Organ", "Disease"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestGeography)) +
  scale_fill_geography(name="Population\nDescriptor") +
  geom_stratum(width=1/3) +
  ylab("Samples") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  theme(legend.position = "bottom")

raceTissueAlluvial <- raceSummary %>% filter(!grepl('cancer', finalSystem)) %>% filter(!grepl('cancer', finalOrgan)) %>% drop_na(c(strictestRace, finalSystem, finalOrgan, finalDisease)) %>%
ggplot(data = .,
       aes(axis1 = finalSystem, axis2 = finalOrgan, axis3 = finalDisease, y = n)) +
  scale_x_discrete(limits = c("Biological system", "Organ", "Disease"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestRace)) +
  scale_fill_brewer(palette="Dark2", name="US Census\nRacial Term") +
  geom_stratum(width=1/3) +
  ylab("Samples") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  theme(legend.position = "bottom")


pdf(file="_tissue_disease_alluvial.pdf")
  geoTissueAlluvial + raceTissueAlluvial +
    plot_layout(nrow=2)
dev.off()

```

## Bonus. How do the descriptor levels relate to each other?

```{r alluvial-descriptors, out.width="100%", include=F, eval=F}

# Let's attempt an alluvial plot and call it quits:
onion <- allSRAFinal %>% count(finalGeography) %>% count(strictGeography) %>% count(strictestGeography) %>% drop_na(c(strictestGeography, strictGeography, finalGeography)) 


%>%
ggplot(data = .,
       aes(axis1 = finalGeography, axis2 = strictGeography, axis3 = strictestGeography, y = n)) +
  scale_x_discrete(limits = c("Original", "Strict", "Strictest"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = finalGeography)) +
  scale_fill_brewer(palette="Set3") +
    geom_stratum(width=1/3) +
  ylab("Samples") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  theme(legend.position = "none")

raceTermsAlluvial <- allSRAFinal %>% count(strictestRace, finalCountry) %>% drop_na(c(strictestRace, finalCountry)) %>%
ggplot(data = .,
       aes(axis1 = finalCountry, axis2 = strictestRace, y = n)) +
  scale_x_discrete(limits = c("SRA Depositor\nCountry", "US Census Racial Term"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestRace)) +
  scale_fill_brewer(palette="Dark2", name="US Census\nRacial Term") +
  geom_stratum(width=1/3) +
  ylab("Samples") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  theme(legend.position = "none")

```