---
title: "Diversity project actual analyses I swear"
layout: "post"
tags: "diversity"
output: html_document
editor_options: 
  chunk_output_type: console
---

We will finally analyse this data properly. 
```{r setup, include=F}
knitr::opts_knit$set(root.dir="~/OneDrive - St Vincent's Institute/Documents/RNA\ Diversity/", echo=T)
# But we also have to do this manually because R studio is stupid: 
setwd(knitr::opts_knit$get("root.dir"))
library(ggplot2)
library(data.table)
library(plyr)
library(reshape)
library(tidyverse)
library(patchwork)
library(ggalluvial)
library(RColorBrewer)
library(ggrepel)
library(ggmosaic)
library(ggforce)
options(device = "quartz")

allSRAFinal <- readRDS("20240704_allSRAFinal.rds")
```

After discussion with Amanda, we'll be using the strictestRace/Geography assignments, see `06_revisiting_population_terms` for more context on what this means.

```{r plotting-prep-1, out.width="100%"}
# First we clean up the plotting labels:
allSRAFinal <- allSRAFinal %>% 
    mutate(strictestRace = gsub("or ", "or\n", strictestRace)) %>%
    mutate(strictestGeography = gsub("and ", "and\n", strictestGeography)) %>%
    mutate(strictestGeography = gsub("^Asia$", "Asia (NOS)", strictestGeography)) %>%
    mutate(hispanic = gsub("^hispanic$", "Hispanic", hispanic)) %>%
      as.data.frame()

# Let's sort out some colour schemes... For Race we simply use Dark2 but we need to keep the mapping consistent, but for Geography we are following Alicia Martin and using a somewhat custom palette that doesn't follow anything standard, and then adding some new colours of our own; I opened a paper in Affinity and grabbed the html values and then added a couple for the new entries (Southeast Asia, Asia NOS):

allSRAFinal$strictestGeography <- factor(allSRAFinal$strictestGeography, levels = c("Subsaharan Africa", "North Africa and\nWestern Asia", "Europe", "South Asia", "Southeast Asia", "East Asia", "Asia (NOS)", "Oceania", "Americas", "Multiple", "Other"))

scale_fill_geography <- function(...){
    ggplot2:::manual_scale('fill',  
        values = setNames(c('#9C8DC3', '#F3D78A', '#DB6968', '#60BC55', '#BCCC45', '#4D97CD', '#04C3C8', '#C69C3A', '#F8984E', '#8B96AD', '#FBC9C4'), levels(allSRAFinal$strictestGeography)), 
    )
}

scale_color_geography <- function(...){
    ggplot2:::manual_scale('color',  
        values = setNames(c('#9C8DC3', '#F3D78A', '#DB6968', '#60BC55', '#BCCC45', '#4D97CD', '#04C3C8', '#C69C3A', '#F8984E', '#8B96AD', '#FBC9C4'), levels(allSRAFinal$strictestGeography)), 
    )
}

raceColours <- brewer.pal(8,"Set2") #8 because we don't plot NAs

# Hispanic label trumps any other label
allSRAFinal$backupRace <- allSRAFinal$strictestRace
allSRAFinal$backupHispanic <- allSRAFinal$hispanic
allSRAFinal <- allSRAFinal %>% mutate(hispanic = na_if(hispanic, "non.hispanic"))
allSRAFinal$strictestRace <- coalesce(allSRAFinal$hispanic, allSRAFinal$strictestRace)
# allSRAFinal <- allSRAFinal %>% mutate(strictestRace = na_if(strictestRace, "non.hispanic"))

allSRAFinal$strictestRace <- factor(allSRAFinal$strictestRace)

scale_color_race <- function(...){
    ggplot2:::manual_scale('color', 
                           values = setNames(raceColours, levels(allSRAFinal$strictestRace)), 
    )
}

scale_fill_race <- function(...){
    ggplot2:::manual_scale('fill', 
        values = setNames(raceColours, levels(allSRAFinal$strictestRace)), 
    )
}

```

We're also going to use some additional info from the World Bank, so let's add those columns in now:

```{r plotting-prep-2}
worldBank <- read.csv("World_Bank_Descriptors.csv")

# Some names need to be fixed, but others are more flexible? I'd rather have USA than United States.
allSRAFinal$finalCountry <- gsub("^UK$", "United Kingdom", allSRAFinal$finalCountry) 
allSRAFinal$finalCountry <- gsub("^Korea$", "South Korea", allSRAFinal$finalCountry)

allSRAFinal$worldBank <- worldBank[match(allSRAFinal$finalCountry, worldBank$Economy),]$Income.group
# allSRAFinal %>% count(finalCountry, worldBank)

# Missing countries are all high income, so we can set them to that manually now that we're happy with the names:
allSRAFinal <- allSRAFinal %>% 
  mutate(worldBank = if_else(is.na(worldBank), "High income", worldBank)) %>%
  mutate(worldBank = if_else(is.na(finalCountry), NA, worldBank ))

# And now the regions they define:
allSRAFinal$worldRegion <- worldBank[match(allSRAFinal$finalCountry, worldBank$Economy),]$Region
# allSRAFinal %>% count(finalCountry, worldRegion)

# Once again we gotta add some manually...
allSRAFinal <- allSRAFinal %>% 
  mutate(worldRegion = if_else(grepl("Russia|Slovakia", finalCountry), "Europe & Central Asia", worldRegion)) %>%
  mutate(worldRegion = if_else(grepl("South Korea|Taiwan", finalCountry), "East Asia & Pacific", worldRegion)) %>%
  mutate(worldRegion = if_else(grepl("USA", finalCountry), "North America", worldRegion)) %>%
  as.data.frame()

# And now we need to clean up some labels so we can plot them better. Good thing is, it's just an ampersand:
allSRAFinal$worldRegion <- gsub("& ", "&\n", allSRAFinal$worldRegion)
```

## 1. Who is being sequenced?

```{r plot-survey-level, out.width="100%"}
geographyProp <- allSRAFinal %>% drop_na(strictestGeography) %>% count(strictestGeography) %>% mutate(freq = n/sum(n))
raceProp <- allSRAFinal %>% drop_na(strictestRace) %>% count(strictestRace) %>% mutate(freq = n/sum(n))

ggplot(geographyProp, aes(x = strictestGeography, y = n, fill = strictestGeography)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Submitted descriptor (ancestral/geographic origin)") +
  xlab("") +
  ylab("Samples") +
  # scale_y_continuous(trans='log10') +
  scale_fill_geography() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none")

geographyProp
sum(geographyProp$n)

ggplot(raceProp, aes(x = strictestRace, y = n, fill=strictestRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Submitted descriptor (US Census term)") +
  xlab("") +
  ylab("Samples") +
  # scale_y_continuous(trans='log10') +
  # scale_fill_brewer(palette="Dark2") +
  scale_fill_race() +
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none")

raceProp
sum(raceProp$n)

# There's 132 samples that do not have any info besides "Not Hispanic", which I think we should treat as NA, and therefore drop. The other option is to change them to race: Other and then move on with our lives. 
allSRAFinal %>% filter(is.na(allSRAFinal$strictestGeography) & is.na(allSRAFinal$strictestRace)) %>% count(ETHNICITY, RACE)
allSRAFinal <- allSRAFinal %>% filter(!is.na(allSRAFinal$strictestGeography) | !is.na(allSRAFinal$strictestRace))

nrow(allSRAFinal[!is.na(allSRAFinal$strictestGeography) | !is.na(allSRAFinal$strictestRace),])
dim(allSRAFinal)
```

All samples accounted for now, at last! 

So let's get a bit more granular, showing n samples per study:

```{r plot-survey-level-2, out.width="100%", echo=F}
sampleGeography <- allSRAFinal %>% count(SRA.Study, strictestGeography) %>% drop_na(strictestGeography)
sampleRace <- allSRAFinal %>% count(SRA.Study, strictestRace) %>% drop_na(strictestRace)
meltSampleGeography <- melt(sampleGeography)
meltSampleRace <- melt(sampleRace)

ggplot(meltSampleGeography, aes(x = strictestGeography, y=value,fill=strictestGeography)) +
  geom_boxplot(width=0.8, outlier.shape = NA) +
  geom_jitter(size=0.8, width=0.2, color="#333333") +
  theme_bw() +
  ggtitle("Submitted descriptor\n(ancestral/geographic origin)") +
  xlab("") +
  ylab("Samples per study") +
  scale_y_continuous(trans='log10') +
  scale_fill_geography() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "none")

ggplot(meltSampleRace, aes(x = strictestRace, y=value,fill=strictestRace)) +
  geom_boxplot(width=0.8, outlier.shape = NA) +
  geom_jitter(size=0.8, width=0.2, color="#333333") +
  theme_bw() +
  ggtitle("Submitted descriptor\n(US Census term)") +
  xlab("") +
  ylab("Samples per study") +
  scale_y_continuous(trans='log10') +
  scale_fill_race() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "none")
```

Let's get some meaningful statistics on this:

First we separate studies on the basis of whether they contain racial or geographic descriptors, but the hispanic usage introduces a new complexity... There's a bunch of studies that now straddle the two categories. Could shift them to race like we did before, but better to just leave them alone for now? I am overthinking.

```{r study-size-statistics}
geographyClean <- allSRAFinal %>% drop_na(strictestGeography) 
raceClean <- allSRAFinal %>% drop_na(strictestRace)

geoSRA <- unique(geographyClean$SRA.Study)
length(geoSRA)
raceSRA <- unique(raceClean$SRA.Study)
length(raceSRA)
length(intersect(geoSRA, raceSRA)) 

allSRAFinal %>% filter(grepl(paste(intersect(geoSRA, raceSRA), collapse="|"), SRA.Study)) %>%
  group_by(SRA.Study) %>%
  summarise(size = n(), studyJoint = n_distinct(SRA.Study)) 

allSRAFinal %>% filter(grepl(paste(intersect(geoSRA, raceSRA), collapse="|"), SRA.Study)) %>% nrow()

allSRAFinal %>% filter(grepl(paste(intersect(geoSRA, raceSRA), collapse="|"), SRA.Study)) %>%
  group_by(SRA.Study) %>%
  count(strictestGeography, strictestRace, finalCountry) %>%
  as.data.frame()

geographyClean %>% group_by(SRA.Study) %>% 
  summarise(size = n(), studyGeo = n_distinct(strictestGeography)) %>%
  summarise(meanSize = mean(size), meanGeo = mean(studyGeo), maxSize = max(size), maxGeo = max(studyGeo))

raceClean %>% group_by(SRA.Study) %>% 
  summarise(size = n(), studyRace = n_distinct(strictestRace)) %>%
  summarise(meanSize = mean(size), meanRace = mean(studyRace), maxSize = max(size), maxRace = max(studyRace))

geographyClean %>% group_by(strictestGeography) %>% 
  summarise(size = n(), studyGeo = n_distinct(SRA.Study)) 

raceClean %>% group_by(strictestRace) %>% 
  summarise(size = n(), studyRace = n_distinct(SRA.Study)) 
```

## 2. And by who? Where is the sequencing happening? 

There are many ways to slice this, but some are harder to see than others, so this section is full of abandoned and rejected plots. But a pair of these plots (race, ancestry) would be great

```{r plot-clean-country, out.width="100%", echo=F}
# ggplot(geographyClean, aes(x = finalCountry, fill = strictestGeography)) +
#   geom_bar(position = position_dodge2(width = 0.9, preserve = "single")) +
#   theme_bw() +
#   ggtitle("SRA Depositor Country") +
#   xlab("") +
#   ylab("Samples") +
#   scale_y_continuous(trans='log10') +
#   scale_fill_geography(name="Population\nDescriptor") +
#   theme(axis.text.x = element_text(angle = 45, hjust=1)) +
#   # theme(legend.position = "bottom")
#   theme(legend.position = "none")
# 
# ggplot(raceClean, aes(x = finalCountry, fill=strictestRace)) +
#   geom_bar(position = position_dodge2(width = 0.9, preserve = "single")) +
#   theme_bw() +
#   ggtitle("SRA Depositor Country") +
#   xlab("") +
#   ylab("Samples") +
#   scale_y_continuous(trans='log10') +
#   scale_fill_race(name="US Census\nTerm)") +
#   theme(axis.text.x = element_text(angle = 45, hjust=1)) +
#   # theme(legend.position = "bottom")
#   theme(legend.position = "none")

# What about stacked bar plots instead of dodge?
geoByCountry <- allSRAFinal %>% drop_na(finalCountry, strictestGeography) %>% count(finalCountry, strictestGeography, worldRegion) %>% melt() %>% arrange(worldRegion) %>% 
ggplot(., aes(x = fct_inorder(finalCountry), y=value, fill = strictestGeography)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  ggtitle("Submitted descriptor\n(ancestral/geographic)") +
  xlab("SRA depositor country") +
  ylab("Number of samples") +
  coord_flip() +
  scale_fill_geography(name = "Population\nDescriptor") + # For some reason this isn't working for setting the name...
  guides(fill=guide_legend(title="Population\nDescriptor", label.position="left")) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.title.align = 1, legend.direction = "vertical", legend.justification = c(1,0), legend.position = c(1,0), legend.background=element_blank())

raceByCountry <- allSRAFinal %>% drop_na(finalCountry, strictestRace) %>% count(finalCountry, strictestRace, worldRegion) %>% melt() %>% arrange(worldRegion) %>% 
ggplot(., aes(x = fct_inorder(finalCountry), y=value, fill = strictestRace)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  ggtitle("Submitted descriptor\n(US Census term)") +
  xlab("SRA depositor country") +
  ylab("Number of samples") +
  coord_flip() +
  scale_fill_race(name="US Census\nTerm") +
  guides(fill=guide_legend(title="US Census\nTerm", label.position = "left")) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.title.align = 1, legend.direction = "vertical", legend.justification = c(1,0), legend.position = c(1,0), legend.background=element_blank())

plotDesign <- "
  12
  12
"
geoByCountry + raceByCountry +
    plot_layout(design = plotDesign)

# And proportional?
allSRAFinal %>% drop_na(finalCountry, strictestGeography) %>% 
ggplot(., aes(x = finalCountry, fill = strictestGeography)) +
  geom_bar(position = "fill") +
  theme_bw() +
  ggtitle("SRA Depositor Country") +
  xlab("") +
  ylab("Sample proportion") +
  # scale_y_continuous(trans='log10') +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

allSRAFinal %>% drop_na(finalCountry, strictestRace) %>%
ggplot(., aes(x = finalCountry, fill = strictestRace)) +
  geom_bar(position = "fill") +
  theme_bw() +
  ggtitle("SRA Depositor Country") +
  xlab("") +
  ylab("Sample proportion") +
  # scale_y_continuous(trans='log10') +
  scale_fill_race(name="US Census\nTerm") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")
```

This is nice but maybe a bit too hard to look at, visually, so let's consider grouping countries by World Bank income status instead:

```{r plot-clean-country-3, out.width="100%"}

# allSRAFinal %>% count(finalCountry, worldBank)

# Some statistics:
allSRAFinal %>% filter(!is.na(worldBank)) %>% count(worldBank) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(finalCountry)) %>% count(finalCountry) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))

allSRAFinal %>% filter(!is.na(strictestGeography)) %>% count(worldBank) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(strictestRace)) %>% count(worldBank) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(hispanic)) %>% count(worldBank) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))

allSRAFinal %>% filter(!is.na(strictestGeography)) %>% count(finalCountry) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(strictestRace)) %>% count(finalCountry) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(hispanic)) %>% count(finalCountry) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
```

Let's attempt some alluvial plots and see if that makes it easier to see?

```{r plot-clean-country-4, out.width="100%", echo=F}
allSRAFinal %>% count(strictestGeography, worldBank) %>% drop_na(c(strictestGeography, worldBank)) %>%
  ggplot(data = .,
         aes(axis1 = worldBank, axis2 = strictestGeography, y = n)) +
    scale_x_discrete(limits = c("World Bank classification", "Population\nDescriptor"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestGeography)) +
    scale_fill_geography(name="Population\nDescriptor") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

allSRAFinal %>% count(strictestRace, worldBank) %>% drop_na(c(strictestRace, worldBank)) %>%
  ggplot(data = .,
         aes(axis1 = worldBank, axis2 = strictestRace, y = n)) +
    scale_x_discrete(limits = c("World Bank classification", "US Census Racial Term"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestRace)) +
    scale_fill_race(name="US Census\nTerm)") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

# These are actually really uninformative, so let's go back to trying to make the ones at the country level and just make them pretty. 
allSRAFinal %>% count(strictestGeography, finalCountry) %>% drop_na(c(strictestGeography, finalCountry)) %>%
  ggplot(data = .,
         aes(axis1 = finalCountry, axis2 = strictestGeography, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nCountry", "Population\nDescriptor"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestGeography)) +
    scale_fill_geography(name="Population\nDescriptor") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

allSRAFinal %>% count(strictestRace, finalCountry) %>% drop_na(c(strictestRace, finalCountry)) %>%
  ggplot(data = .,
         aes(axis1 = finalCountry, axis2 = strictestRace, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nCountry", "US Census Racial Term"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestRace)) +
    scale_fill_race(name="US Census\nTerm)") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")
```

```{r plot-clean-country-5}
allSRAFinal %>% filter(!is.na(strictestGeography)) %>% count(worldRegion) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(strictestRace)) %>% count(worldRegion) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% count(worldRegion) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
# allSRAFinal %>% count(finalCountry, worldRegion)
```

```{r plot-clean-country-6, out.width="100%", echo=F}
allSRAFinal %>% count(strictestGeography, worldRegion) %>% drop_na(c(strictestGeography, worldRegion)) %>%
  ggplot(data = .,
         aes(axis1 = worldRegion, axis2 = strictestGeography, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nRegion", "Population\nDescriptor"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestGeography)) +
    scale_fill_geography(name="Population\nDescriptor") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

allSRAFinal %>% count(strictestRace, worldRegion) %>% drop_na(c(strictestRace, worldRegion)) %>%
  ggplot(data = .,
         aes(axis1 = worldRegion, axis2 = strictestRace, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nRegion", "US Census Racial Term"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestRace)) +
    scale_fill_race(name="US Census\nTerm)") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

### And finally... does it have any kind of description??
allSRAFinal <- allSRAFinal %>% 
  mutate(hasDescriptor = if_else(is.na(strictestRace), "Geographic", "Racial"))

# Useless at this resolution
allSRAFinal %>% count(finalCountry, hasDescriptor) %>%
  ggplot(data = .,
         aes(axis1 = finalCountry, axis2 = hasDescriptor, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nCountry", "Descriptor\nType"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = hasDescriptor)) +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

allSRAFinal %>% count(worldRegion, hasDescriptor) %>%
  ggplot(data = .,
         aes(axis1 = worldRegion, axis2 = hasDescriptor, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nRegion", "Descriptor\nType"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = hasDescriptor)) +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

# What about some bar charts instead...

# What about stacked bar plots instead of dodge?
allSRAFinal %>% drop_na(worldRegion, strictestGeography) %>% count(worldRegion, strictestGeography) %>% melt() %>%
ggplot(., aes(x = worldRegion, y=value, fill = strictestGeography)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  ggtitle("SRA Depositor Country") +
  xlab("") +
  ylab("Samples") +
  # scale_y_continuous(trans='log10') +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

allSRAFinal %>% drop_na(worldRegion, strictestRace) %>% count(worldRegion, strictestRace) %>% melt() %>%
ggplot(., aes(x = worldRegion, y=value, fill = strictestRace)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  ggtitle("SRA Depositor Country") +
  xlab("") +
  ylab("Samples") +
  coord_flip() +
  # scale_y_continuous(trans='log10') +
  scale_fill_race(name="US Census\nTerm") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")



# And proportional?
allSRAFinal %>% drop_na(worldRegion, strictestGeography) %>% 
ggplot(., aes(x = worldRegion, fill = strictestGeography)) +
  geom_bar(position = "fill") +
  theme_bw() +
  ggtitle("SRA Depositor Country") +
  xlab("") +
  ylab("Sample proportion") +
  # scale_y_continuous(trans='log10') +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

allSRAFinal %>% drop_na(worldRegion, strictestRace) %>%
ggplot(., aes(x = worldRegion, fill = strictestRace)) +
  geom_bar(position = "fill") +
  theme_bw() +
  ggtitle("SRA Depositor Country") +
  xlab("") +
  ylab("Sample proportion") +
  # scale_y_continuous(trans='log10') +
  scale_fill_race(name="US Census\nTerm") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")




```

And now, some statistics from the figures above:
  
```{r consolidated-stats}
allSRAFinal %>% count(finalCountry, hasDescriptor) %>% 
  group_by(finalCountry) %>%
  mutate(percent = 100 * n/sum(n)) %>%
  ungroup %>%
  as.data.frame()
                
allSRAFinal %>% filter(!grepl('USA', finalCountry)) %>% 
  count(hasDescriptor) %>% 
  mutate(freq = n/sum(n))

allSRAFinal %>% count(hasDescriptor) %>% 
  mutate(freq = n/sum(n))

# We also want to know the mean sample size of studies in the USA vs those not in the USA, for funsies
usaSizes <- allSRAFinal %>% filter(grepl('USA', finalCountry)) %>%
  group_by(SRA.Study) %>%
  summarise(n = n())
mean(usaSizes$n)

nousaSizes <- allSRAFinal %>% filter(!grepl('USA', finalCountry)) %>%
  group_by(SRA.Study) %>%
  summarise(n = n())
mean(nousaSizes$n)

t.test(usaSizes$n, nousaSizes$n)
```


## 3. What tissues are they sequencing?


```{r first-pass-tissue, out.width="100%"}
# First we focus on population descriptors:
bigGeographySummary <- allSRAFinal %>% count(SRA.Study, strictestGeography, finalSystem, finalOrgan, finalDisease)
bigRaceSummary <- allSRAFinal %>% count(SRA.Study, strictestRace, finalSystem, finalOrgan, finalDisease)
geographySummary <- allSRAFinal %>% count(strictestGeography, finalSystem, finalOrgan, finalDisease)
raceSummary <- allSRAFinal %>% count(strictestRace, finalSystem, finalOrgan, finalDisease)

meltGeography <- melt(geographySummary)
meltRace <- melt(raceSummary)

meltGeography %>% drop_na(c(finalOrgan, strictestGeography)) %>%
  ggplot(., aes(x = finalOrgan, y = value, fill = strictestGeography)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported organ") +
  xlab("") +
  ylab("Samples") +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "bottom")

meltGeography %>% drop_na(c(finalDisease, strictestGeography)) %>%
  ggplot(., aes(x = finalDisease, y = value, fill = strictestGeography)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported disease") +
  xlab("") +
  ylab("Samples") +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

meltRace %>% drop_na(c(finalOrgan, strictestRace)) %>%
  ggplot(., aes(x = finalOrgan, y = value, fill = strictestRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported organ") +
  xlab("") +
  ylab("Samples") +
  scale_fill_race(name="US Census\nTerm)") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "bottom")

meltRace %>% drop_na(c(finalDisease, strictestRace)) %>%
  ggplot(., aes(x = finalDisease, y = value, fill = strictestRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported disease") +
  xlab("") +
  ylab("Samples") +
  scale_fill_race(name="US Census\nTerm)") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

```

And now a bit of stats here too...

```{r first-pass-tissue-2, out.width="100%"}
# First, is blood really more diverse than anyhting else? 

bloodPropGeo <- allSRAFinal %>% filter(grepl("blood", finalOrgan) & !is.na(strictestGeography)) %>% 
  count(strictestGeography) %>% 
  mutate(freq = n/sum(n))

nobloodPropGeo <- allSRAFinal %>% filter(!grepl("blood", finalOrgan) & !is.na(strictestGeography)) %>% 
  count(strictestGeography) %>% 
  mutate(freq = n/sum(n))

bloodPropRace <- allSRAFinal %>% filter(grepl("blood", finalOrgan) & !is.na(strictestRace)) %>% 
  count(strictestRace) %>% 
  mutate(freq = n/sum(n))
  
nobloodPropRace <- allSRAFinal %>% filter(!grepl("blood", finalOrgan) & !is.na(strictestRace)) %>% 
  count(strictestRace) %>% 
  mutate(freq = n/sum(n))

geoBlood <- full_join(bloodPropGeo, nobloodPropGeo, by="strictestGeography", suffix=c("Blood", "NoBlood")) %>%
  full_join(., geographyProp, by="strictestGeography") %>%
  mutate(across(where(is.numeric), ~replace(., is.na(.), 0)))

raceBlood <- full_join(bloodPropRace, nobloodPropRace, by="strictestRace", suffix=c("Blood", "NoBlood")) %>%
  full_join(., raceProp, by="strictestRace") %>%
  mutate(across(where(is.numeric), ~replace(., is.na(.), 0)))

chisq.test(geoBlood[,c(2,6)])
chisq.test(raceBlood[,c(2,6)])

# Both are significant but blood is clearly a lot more diverse
geoBlood
raceBlood
```
What about diseases? Are there associations between place and type?

From here onwards, we set we're gonna set `cancer` in either `finalOrgan` or `finalSystem` when we also consider `finalDisease`, because otherwise it's a bit weird


```{r first-pass-disease-1, out.width="100%"}

```



## 4. Attempting better ways to look at some of the relationships between classes:

For these (and only for these!), we're gonna set `cancer` in either `finalOrgan` or `finalSystem` to NA, because otherwise the data gets messy. Worth discussing what to do in general/above. 

```{r attempt-alluvial-1, out.width="100%"}
allSRAFinal %>% count(strictestGeography, worldRegion, finalOrgan, finalDisease) %>% filter(!grepl('cancer', finalOrgan)) %>% drop_na(c(strictestGeography, worldRegion, finalOrgan, finalDisease)) %>%
  ggplot(data = .,
       aes(axis1 = worldRegion, axis2 = finalOrgan, axis3 = finalDisease, y = n)) +
  scale_x_discrete(limits = c("SRA Depositor Region", "Organ", "Disease group"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestGeography)) +
  scale_fill_geography(name="Population\nDescriptor") +
  geom_stratum(width=1/3) +
  ylab("Samples") +
  # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
  theme_minimal() +
  theme(legend.position = "bottom")

allSRAFinal %>% count(strictestRace, worldRegion, finalOrgan, finalDisease) %>% filter(!grepl('cancer', finalOrgan)) %>% drop_na(c(strictestRace, worldRegion, finalOrgan, finalDisease)) %>%
  ggplot(data = .,
       aes(axis1 = worldRegion, axis2 = finalOrgan, axis3 = finalDisease, y = n)) +
  scale_x_discrete(limits = c("SRA Depositor\nRegion", "Organ", "Disease"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestRace)) +
  scale_fill_race(name="US Census\nTerm") +
  geom_stratum(width=1/3) +
  ylab("Samples") +
  # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
  theme_minimal() +
  theme(legend.position = "bottom")

# allSRAFinal %>% count(strictestGeography, worldRegion, finalSystem, finalDisease) %>% filter(!grepl('cancer', finalSystem)) %>% drop_na(c(strictestGeography, worldRegion, finalSystem, finalDisease)) %>%
#   ggplot(data = .,
#        aes(axis1 = worldRegion, axis2 = finalSystem, axis3 = finalDisease, y = n)) +
#   scale_x_discrete(limits = c("SRA Depositor\nRegion", "Biological\nSystem", "Disease"), expand = c(.2, .05)) +
#   geom_alluvium(aes(fill = strictestGeography)) +
#   scale_fill_geography(name="Population\nDescriptor") +
#   geom_stratum(width=1/3) +
#   ylab("Samples") +
#   # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
#   geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
#   theme_minimal() +
#   theme(legend.position = "bottom")
# 
# allSRAFinal %>% count(strictestRace, worldRegion, finalSystem, finalDisease) %>% filter(!grepl('cancer', finalSystem)) %>% drop_na(c(strictestRace, worldRegion, finalSystem, finalDisease)) %>%
#   ggplot(data = .,
#        aes(axis1 = worldRegion, axis2 = finalSystem, axis3 = finalDisease, y = n)) +
#   scale_x_discrete(limits = c("SRA Depositor\nRegion", "Biological\nSystem", "Disease"), expand = c(.2, .05)) +
#   geom_alluvium(aes(fill = strictestRace)) +
#   scale_fill_race(name="US Census\nTerm") +
#   geom_stratum(width=1/3) +
#   ylab("Samples") +
#   # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
#   geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
#   theme_minimal() +
#   theme(legend.position = "bottom")

```

Alluvial plots are too confusing, so let's consider other approaches... mosaic plots?

```{r attempt-mosaic-1}
allSRAFinal %>% filter(!grepl('cancer', finalOrgan)) %>% drop_na(finalDisease) %>% drop_na(strictestGeography) %>%
ggplot(.) +
  geom_mosaic(aes(x = product(finalDisease), fill=finalOrgan)) +
  facet_grid(~strictestGeography)
```

Yeah no. A different kind of alluvial plot, easier to manipulate?

```{r attempt-alluvial-2}
diseaseTry <- allSRAFinal %>% drop_na(strictestGeography, finalDisease, finalOrgan) %>% count(strictestGeography, finalDisease, finalOrgan)
diseaseTry <- gather_set_data(diseaseTry, c(2, 3, 1))
head(diseaseTry)

ggplot(diseaseTry, aes(x, id = id, split = y, value = n)) +
  geom_parallel_sets(aes(fill = strictestGeography), alpha = 0.3, axis.width = 0.1) +
  geom_parallel_sets_axes(axis.width = 0.1) +
  geom_parallel_sets_labels(colour = 'white') +
  scale_fill_geography() +
  theme_bw()
```

Also no. Bubble plot?

```{r attempt-bubble-1}
allSRAFinal %>% drop_na(strictestGeography, finalDisease, finalOrgan) %>%
ggplot(., aes(x=finalDisease, y=finalOrgan)) +
  geom_count() +
  facet_grid(~strictestGeography) +
  theme_bw()

```


## Bonus. How do the descriptor levels relate to each other?

```{r alluvial-descriptors, out.width="100%", include=F, eval=F}

# Let's attempt an alluvial plot and call it quits:
allSRAFinal %>% count(finalGeography, strictestGeography, finalRace, strictestRace) %>%
ggplot(data = .,
       aes(axis1 = finalGeography, axis2 = strictestGeography, axis3 = finalRace, axis4 = strictestRace, y = n)) +
  scale_x_discrete(limits = c("Original", "Strict", "Strictest"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = finalGeography)) +
  # scale_fill_brewer(palette="Set3") +
    geom_stratum(width=1/3) +
  ylab("Samples") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  theme(legend.position = "none")
```
