---
title: "Diversity project actual analyses I swear"
layout: "post"
tags: "diversity"
output: html_document
editor_options: 
  chunk_output_type: console
---

We will finally analyse this data properly. 
```{r setup, include=F}
knitr::opts_knit$set(root.dir="~/OneDrive - St Vincent's Institute/Documents/RNA\ Diversity/", echo=T)
# But we also have to do this manually because R studio is stupid: 
setwd(knitr::opts_knit$get("root.dir"))
library(ggplot2)
library(data.table)
library(plyr)
library(reshape)
library(tidyverse)
library(patchwork)
library(ggalluvial)
library(RColorBrewer)
library(ggrepel)
options(device = "quartz")

allSRAFinal <- readRDS("20240704_allSRAFinal.rds")
```

After discussion with Amanda, we'll be using the strictestRace/Geography assignments, see `06_revisiting_population_terms` for more context on what this means.

```{r plotting-prep, out.width="100%"}
# First we clean up the plotting labels:
allSRAFinal <- allSRAFinal %>% 
    mutate(strictestRace = gsub("or ", "or\n", strictestRace)) %>%
    mutate(strictestGeography = gsub("and ", "and\n", strictestGeography)) %>%
    mutate(strictestGeography = gsub("^Asia$", "Asia (NOS)", strictestGeography)) %>%
    mutate(hispanic = gsub("^hispanic$", "Hispanic", hispanic)) %>%
      as.data.frame()

# Let's sort out some colour schemes... For Race we simply use Dark2 but we need to keep the mapping consistent, but for Geography we are following Alicia Martin and using a somewhat custom palette that doesn't follow anything standard, and then adding some new colours of our own; I opened a paper in Affinity and grabbed the html values and then added a couple for the new entries (Southeast Asia, Asia NOS):

allSRAFinal$strictestGeography <- factor(allSRAFinal$strictestGeography, levels = c("Subsaharan Africa", "North Africa and\nWestern Asia", "Europe", "South Asia", "Southeast Asia", "East Asia", "Asia (NOS)", "Oceania", "Americas", "Multiple", "Other"))

scale_fill_geography <- function(...){
    ggplot2:::manual_scale('fill',  
        values = setNames(c('#9C8DC3', '#F3D78A', '#DB6968', '#60BC55', '#BCCC45', '#4D97CD', '#04C3C8', '#C69C3A', '#F8984E', '#8B96AD', '#FBC9C4'), levels(allSRAFinal$strictestGeography)), 
    )
}

scale_color_geography <- function(...){
    ggplot2:::manual_scale('color',  
        values = setNames(c('#9C8DC3', '#F3D78A', '#DB6968', '#60BC55', '#BCCC45', '#4D97CD', '#04C3C8', '#C69C3A', '#F8984E', '#8B96AD', '#FBC9C4'), levels(allSRAFinal$strictestGeography)), 
    )
}

raceColours <- brewer.pal(7,"Set2") #8 because we don't plot NAs

# allSRAFinal$strictestRace <- coalesce(allSRAFinal$strictestRace, allSRAFinal$hispanic)
# allSRAFinal <- allSRAFinal %>% mutate(strictestRace = na_if(strictestRace, "non.hispanic"))

allSRAFinal$strictestRace <- factor(allSRAFinal$strictestRace)

scale_color_race <- function(...){
    ggplot2:::manual_scale('color', 
                           values = setNames(raceColours, levels(allSRAFinal$strictestRace)), 
    )
}

scale_fill_race <- function(...){
    ggplot2:::manual_scale('fill', 
        values = setNames(raceColours, levels(allSRAFinal$strictestRace)), 
    )
}

```

## 1. Who is being sequenced?

```{r plot-survey-level, out.width="100%"}
geographyClean <- allSRAFinal %>% drop_na(strictestGeography) %>% count(strictestGeography) %>% mutate(freq = n/sum(n))
raceClean <- allSRAFinal %>% drop_na(strictestRace) %>% count(strictestRace) %>% mutate(freq = n/sum(n))

ggplot(geographyClean, aes(x = strictestGeography, y = n, fill = strictestGeography)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Submitted descriptor (ancestral/geographic origin)") +
  xlab("") +
  ylab("Samples") +
  # scale_y_continuous(trans='log10') +
  scale_fill_geography() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none")

geographyClean
sum(geographyClean$n)

ggplot(raceClean, aes(x = strictestRace, y = n, fill=strictestRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Submitted descriptor (US Census term)") +
  xlab("") +
  ylab("Samples") +
  # scale_y_continuous(trans='log10') +
  # scale_fill_brewer(palette="Dark2") +
  scale_fill_race() +
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none")

raceClean
sum(raceClean$n)
allSRAFinal %>% drop_na(strictestRace) %>% filter(!grepl("Hispanic", strictestRace)) %>% count(strictestRace) %>% mutate(freq = n/sum(n))

nrow(allSRAFinal[!is.na(allSRAFinal$strictestGeography) | !is.na(allSRAFinal$strictestRace),])
dim(allSRAFinal)

# There's a discrepancy... but hispanic is now coded in strictestRace, so what is going on?
nrow(allSRAFinal) - nrow(allSRAFinal[!is.na(allSRAFinal$strictestGeography) | !is.na(allSRAFinal$strictestRace),])
allSRAFinal[is.na(allSRAFinal$strictestGeography) & is.na(allSRAFinal$strictestRace),] %>% count(hispanic)
```

All samples accounted for now. Pending a decision on the hispanic samples, then, the final numbers are:

With hispanic samples:
```{r}
nrow(allSRAFinal)
```

Without hispanic samples:

```{r}
nrow(allSRAFinal[!is.na(allSRAFinal$strictestGeography) | !is.na(allSRAFinal$strictestRace),])
```

And now a bit more granular, showing n samples per study:

```{r plot-survey-level-2, out.width="100%", echo=F}
sampleGeography <- allSRAFinal %>% count(SRA.Study, strictestGeography) %>% drop_na(strictestGeography)
sampleRace <- allSRAFinal %>% count(SRA.Study, strictestRace) %>% drop_na(strictestRace)
meltSampleGeography <- melt(sampleGeography)
meltSampleRace <- melt(sampleRace)

ggplot(meltSampleGeography, aes(x = strictestGeography, y=value,fill=strictestGeography)) +
  geom_boxplot(width=0.8, outlier.shape = NA) +
  geom_jitter(size=0.8, width=0.2, color="#333333") +
  theme_bw() +
  ggtitle("Submitted descriptor\n(ancestral/geographic origin)") +
  xlab("") +
  ylab("Samples per study") +
  scale_y_continuous(trans='log10') +
  scale_fill_geography() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "none")

ggplot(meltSampleRace, aes(x = strictestRace, y=value,fill=strictestRace)) +
  geom_boxplot(width=0.8, outlier.shape = NA) +
  geom_jitter(size=0.8, width=0.2, color="#333333") +
  theme_bw() +
  ggtitle("Submitted descriptor\n(US Census term)") +
  xlab("") +
  ylab("Samples per study") +
  scale_y_continuous(trans='log10') +
  scale_fill_race() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "none")
```

Let's get some meaningful statistics on this:

```{r}
# Since no study contains both geo and race info, we can separate them out and make our lives easier:
geographyClean <- allSRAFinal %>% drop_na(strictestGeography) 
raceClean <- allSRAFinal %>% drop_na(strictestRace)
hispanicClean <- allSRAFinal %>% drop_na(hispanic)

length(unique(allSRAFinal$SRA.Study))
length(unique(geographyClean$SRA.Study))
length(unique(raceClean$SRA.Study))
length(unique(hispanicClean$SRA.Study))

geoSRA <- unique(geographyClean$SRA.Study)
raceSRA <- unique(raceClean$SRA.Study)
hispanicSRA <- unique(hispanicClean$SRA.Study)
length(unique(c(geoSRA, raceSRA, hispanicSRA))) 

geographyClean %>% group_by(SRA.Study) %>% 
  summarise(size = n(), studyGeo = n_distinct(strictestGeography)) %>%
  summarise(meanSize = mean(size), meanGeo = mean(studyGeo), maxSize = max(size), maxGeo = max(studyGeo))

raceClean %>% group_by(SRA.Study) %>% 
  summarise(size = n(), studyRace = n_distinct(strictestRace)) %>%
  summarise(meanSize = mean(size), meanRace = mean(studyRace), maxSize = max(size), maxRace = max(studyRace))

hispanicClean %>% group_by(SRA.Study) %>% 
  summarise(size = n(), studyHis = n_distinct(hispanic)) %>%
  summarise(meanSize = mean(size), meanHis = mean(studyHis), maxSize = max(size), maxHis = max(studyHis))

geographyClean %>% group_by(strictestGeography) %>% 
  summarise(size = n(), studyGeo = n_distinct(SRA.Study)) 

raceClean %>% group_by(strictestRace) %>% 
  summarise(size = n(), studyRace = n_distinct(SRA.Study)) 

hispanicClean %>% group_by(hispanic) %>% 
  summarise(size = n(), studyHis = n_distinct(SRA.Study)) 

```

## 2. And by who? Where is the sequencing happening?

```{r plot-clean-country, out.width="100%", echo=F}
ggplot(geographyClean, aes(x = finalCountry, fill = strictestGeography)) +
  geom_bar(position = position_dodge2(width = 0.9, preserve = "single")) +
  theme_bw() +
  ggtitle("SRA Depositor Country") +
  xlab("") +
  ylab("Samples") +
  scale_y_continuous(trans='log10') +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

ggplot(raceClean, aes(x = finalCountry, fill=strictestRace)) +
  geom_bar(position = position_dodge2(width = 0.9, preserve = "single")) +
  theme_bw() +
  ggtitle("SRA Depositor Country") +
  xlab("") +
  ylab("Samples") +
  scale_y_continuous(trans='log10') +
  scale_fill_race(name="US Census\nTerm)") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")
```

This is nice but maybe a bit too hard to look at, visually, so let's consider grouping countries by World Bank income status instead:

```{r plot-clean-country-3, out.width="100%"}
worldBank <- read.csv("World_Bank_Descriptors.csv")

# Some names need to be fixed, but others are more flexible? I'd rather have USA than United States.
allSRAFinal$finalCountry <- gsub("^UK$", "United Kingdom", allSRAFinal$finalCountry) 
allSRAFinal$finalCountry <- gsub("^Korea$", "South Korea", allSRAFinal$finalCountry)

allSRAFinal$worldBank <- worldBank[match(allSRAFinal$finalCountry, worldBank$Economy),]$Income.group
# allSRAFinal %>% count(finalCountry, worldBank)

# Missing countries are all high income, so we can set them to that manually now that we're happy with the names:
allSRAFinal <- allSRAFinal %>% 
  mutate(worldBank = if_else(is.na(worldBank), "High income", worldBank)) %>%
  mutate(worldBank = if_else(is.na(finalCountry), NA, worldBank ))

# allSRAFinal %>% count(finalCountry, worldBank)

# Some statistics:
allSRAFinal %>% filter(!is.na(worldBank)) %>% count(worldBank) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(finalCountry)) %>% count(finalCountry) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))

allSRAFinal %>% filter(!is.na(strictestGeography)) %>% count(worldBank) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(strictestRace)) %>% count(worldBank) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(hispanic)) %>% count(worldBank) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))

allSRAFinal %>% filter(!is.na(strictestGeography)) %>% count(finalCountry) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(strictestRace)) %>% count(finalCountry) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(hispanic)) %>% count(finalCountry) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
```

Let's attempt some alluvial plot and call it quits:

```{r plot-clean-country-4, out.width="100%", echo=F}
allSRAFinal %>% count(strictestGeography, worldBank) %>% drop_na(c(strictestGeography, worldBank)) %>%
  ggplot(data = .,
         aes(axis1 = worldBank, axis2 = strictestGeography, y = n)) +
    scale_x_discrete(limits = c("World Bank classification", "Population\nDescriptor"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestGeography)) +
    scale_fill_geography(name="Population\nDescriptor") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

allSRAFinal %>% count(strictestRace, worldBank) %>% drop_na(c(strictestRace, worldBank)) %>%
  ggplot(data = .,
         aes(axis1 = worldBank, axis2 = strictestRace, y = n)) +
    scale_x_discrete(limits = c("World Bank classification", "US Census Racial Term"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestRace)) +
    scale_fill_race(name="US Census\nTerm)") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

# These are actually really uninformative, so let's go back to trying to make the ones at the country level and just make them pretty. 
allSRAFinal %>% count(strictestGeography, finalCountry) %>% drop_na(c(strictestGeography, finalCountry)) %>%
  ggplot(data = .,
         aes(axis1 = finalCountry, axis2 = strictestGeography, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nCountry", "Population\nDescriptor"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestGeography)) +
    scale_fill_geography(name="Population\nDescriptor") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

allSRAFinal %>% count(strictestRace, finalCountry) %>% drop_na(c(strictestRace, finalCountry)) %>%
  ggplot(data = .,
         aes(axis1 = finalCountry, axis2 = strictestRace, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nCountry", "US Census Racial Term"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestRace)) +
    scale_fill_race(name="US Census\nTerm)") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

# This is also not great... what if we grouped them by world bank economic group?
allSRAFinal$worldRegion <- worldBank[match(allSRAFinal$finalCountry, worldBank$Economy),]$Region
# allSRAFinal %>% count(finalCountry, worldRegion)

# Once again we gotta add some manually...
allSRAFinal <- allSRAFinal %>% 
  mutate(worldRegion = if_else(grepl("Russia|Slovakia", finalCountry), "Europe & Central Asia", worldRegion)) %>%
  mutate(worldRegion = if_else(grepl("South Korea|Taiwan", finalCountry), "East Asia & Pacific", worldRegion)) %>%
  mutate(worldRegion = if_else(grepl("USA", finalCountry), "North America", worldRegion)) %>%
  as.data.frame()

allSRAFinal %>% filter(!is.na(strictestGeography)) %>% count(worldRegion) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% filter(!is.na(strictestRace)) %>% count(worldRegion) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
allSRAFinal %>% count(worldRegion) %>% mutate(freq = n/sum(n)) %>% arrange(desc(freq))
# allSRAFinal %>% count(finalCountry, worldRegion)
allSRAFinal %>% count(strictestGeography, worldRegion) %>% drop_na(c(strictestGeography, worldRegion)) %>%
  ggplot(data = .,
         aes(axis1 = worldRegion, axis2 = strictestGeography, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nRegion", "Population\nDescriptor"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestGeography)) +
    scale_fill_geography(name="Population\nDescriptor") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

allSRAFinal %>% count(strictestRace, worldRegion) %>% drop_na(c(strictestRace, worldRegion)) %>%
  ggplot(data = .,
         aes(axis1 = worldRegion, axis2 = strictestRace, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nRegion", "US Census Racial Term"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = strictestRace)) +
    scale_fill_race(name="US Census\nTerm)") +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")

### And finally... does it have any kind of description??
allSRAFinal <- allSRAFinal %>% 
  mutate(hasDescriptor = if_else(is.na(finalRace), "Geographic", "Racial"))

# Useless at this resolution
allSRAFinal %>% count(finalCountry, hasDescriptor) %>%
  ggplot(data = .,
         aes(axis1 = finalCountry, axis2 = hasDescriptor, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nCountry", "Descriptor\nType"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = hasDescriptor)) +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")


allSRAFinal %>% count(worldRegion, hasDescriptor) %>%
  ggplot(data = .,
         aes(axis1 = worldRegion, axis2 = hasDescriptor, y = n)) +
    scale_x_discrete(limits = c("SRA Depositor\nRegion", "Descriptor\nType"), expand = c(.2, .05)) +
    geom_alluvium(aes(fill = hasDescriptor)) +
    geom_stratum(width=1/3) +
    ylab("Samples") +
    # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
    theme_minimal() +
    theme(legend.position = "none")
```

And now, some statistics from the figures above:
  
```{r consolidated-stats}
allSRAFinal %>% count(finalCountry, hasDescriptor) %>% 
  group_by(finalCountry) %>%
  mutate(percent = 100 * n/sum(n)) %>%
  ungroup %>%
  as.data.frame()
                
allSRAFinal %>% filter(!grepl('USA', finalCountry)) %>% 
  count(hasDescriptor) %>% 
  mutate(freq = n/sum(n))

allSRAFinal %>% filter(!grepl('Hispanic', strictestRace)) %>% 
  count(hasDescriptor) %>% 
  mutate(freq = n/sum(n))

allSRAFinal %>% count(hasDescriptor) %>% 
  mutate(freq = n/sum(n))

# We also want to know the mean sample size of studies in the USA vs those not in the USA, for funsies
usaSizes <- allSRAFinal %>% filter(grepl('USA', finalCountry)) %>%
  group_by(SRA.Study) %>%
  summarise(n = n())

nousaSizes <- allSRAFinal %>% filter(!grepl('USA', finalCountry)) %>%
  group_by(SRA.Study) %>%
  summarise(n = n())

t.test(usaSizes$n, nousaSizes$n)
```


## 3. What tissues are they sequencing?


```{r first-pass-tissue, out.width="100%"}
# First we focus on population descriptors:
bigGeographySummary <- allSRAFinal %>% count(SRA.Study, strictestGeography, finalSystem, finalOrgan, finalDisease)
bigRaceSummary <- allSRAFinal %>% count(SRA.Study, strictestRace, finalSystem, finalOrgan, finalDisease)
geographySummary <- allSRAFinal %>% count(strictestGeography, finalSystem, finalOrgan, finalDisease)
raceSummary <- allSRAFinal %>% count(strictestRace, finalSystem, finalOrgan, finalDisease)

meltGeography <- melt(geographySummary)
meltRace <- melt(raceSummary)

meltGeography %>% drop_na(c(finalSystem, strictestGeography)) %>%
  ggplot(., aes(x = finalSystem, y = value, fill = strictestGeography)) +
  geom_bar(stat="identity") +
  ggtitle("Reported biological system") +
  xlab("") +
  ylab("Samples") +
  theme_bw() +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

meltGeography %>% drop_na(c(finalOrgan, strictestGeography)) %>%
  ggplot(., aes(x = finalOrgan, y = value, fill = strictestGeography)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported organ") +
  xlab("") +
  ylab("Samples") +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "bottom")

meltGeography %>% drop_na(c(finalDisease, strictestGeography)) %>%
  ggplot(., aes(x = finalDisease, y = value, fill = strictestGeography)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported disease") +
  xlab("") +
  ylab("Samples") +
  scale_fill_geography(name="Population\nDescriptor") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")


meltRace %>% drop_na(c(finalSystem, strictestRace)) %>%
  ggplot(., aes(x = finalSystem, y = value, fill = strictestRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported biological system") +
  xlab("") +
  ylab("Samples") +
  scale_fill_race(name="US Census\nTerm)") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

meltRace %>% drop_na(c(finalOrgan, strictestRace)) %>%
  ggplot(., aes(x = finalOrgan, y = value, fill = strictestRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported organ") +
  xlab("") +
  ylab("Samples") +
  scale_fill_race(name="US Census\nTerm)") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position = "bottom")

meltRace %>% drop_na(c(finalDisease, strictestRace)) %>%
  ggplot(., aes(x = finalDisease, y = value, fill = strictestRace)) +
  geom_bar(stat="identity") +
  theme_bw() +
  ggtitle("Reported disease") +
  xlab("") +
  ylab("Samples") +
  scale_fill_race(name="US Census\nTerm)") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  # theme(legend.position = "bottom")
  theme(legend.position = "none")

```

```{r arrange-plots, out.width="100%"}

  # allGeo + geoCountry + geoStudy + geoSystem + geoOrgan + geoDisease +
  #   plot_layout(nrow=)
  # 
  # allRace + raceCountry + raceStudy + raceSystem + raceOrgan + raceDisease +
  #   plot_layout(nrow=2)
  # 
  # allGeo + geoCountry + allRace + raceCountry +
  #   plot_layout(nrow=2, widths = c(1, 2))

```


## 4. Attempting alluvial plots...

For these (and only for these!), we're gonna set `cancer` in either `finalOrgan` or `finalSystem` to NA, because otherwise the data gets messy. Worth discussing what to do in general/above. 

```{r attempt-alluvial-1, out.width="100%"}
allSRAFinal %>% count(strictestGeography, worldRegion, finalOrgan, finalDisease) %>% filter(!grepl('cancer', finalOrgan)) %>% drop_na(c(strictestGeography, worldRegion, finalOrgan, finalDisease)) %>%
  ggplot(data = .,
       aes(axis1 = worldRegion, axis2 = finalOrgan, axis3 = finalDisease, y = n)) +
  scale_x_discrete(limits = c("SRA Depositor\nRegion", "Organ", "Disease"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestGeography)) +
  scale_fill_geography(name="Population\nDescriptor") +
  geom_stratum(width=1/3) +
  ylab("Samples") +
  # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
  theme_minimal() +
  theme(legend.position = "bottom")

allSRAFinal %>% count(strictestRace, worldRegion, finalOrgan, finalDisease) %>% filter(!grepl('cancer', finalOrgan)) %>% drop_na(c(strictestRace, worldRegion, finalOrgan, finalDisease)) %>%
  ggplot(data = .,
       aes(axis1 = worldRegion, axis2 = finalOrgan, axis3 = finalDisease, y = n)) +
  scale_x_discrete(limits = c("SRA Depositor\nRegion", "Organ", "Disease"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestRace)) +
  scale_fill_race(name="US Census\nTerm") +
  geom_stratum(width=1/3) +
  ylab("Samples") +
  # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
  theme_minimal() +
  theme(legend.position = "bottom")

allSRAFinal %>% count(strictestGeography, worldRegion, finalSystem, finalDisease) %>% filter(!grepl('cancer', finalSystem)) %>% drop_na(c(strictestGeography, worldRegion, finalSystem, finalDisease)) %>%
  ggplot(data = .,
       aes(axis1 = worldRegion, axis2 = finalSystem, axis3 = finalDisease, y = n)) +
  scale_x_discrete(limits = c("SRA Depositor\nRegion", "Biological\nSystem", "Disease"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestGeography)) +
  scale_fill_geography(name="Population\nDescriptor") +
  geom_stratum(width=1/3) +
  ylab("Samples") +
  # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
  theme_minimal() +
  theme(legend.position = "bottom")

allSRAFinal %>% count(strictestRace, worldRegion, finalSystem, finalDisease) %>% filter(!grepl('cancer', finalSystem)) %>% drop_na(c(strictestRace, worldRegion, finalSystem, finalDisease)) %>%
  ggplot(data = .,
       aes(axis1 = worldRegion, axis2 = finalSystem, axis3 = finalDisease, y = n)) +
  scale_x_discrete(limits = c("SRA Depositor\nRegion", "Biological\nSystem", "Disease"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestRace)) +
  scale_fill_race(name="US Census\nTerm") +
  geom_stratum(width=1/3) +
  ylab("Samples") +
  # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 4, direction = "y", nudge_x = .5) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

## Bonus. How do the descriptor levels relate to each other?

```{r alluvial-descriptors, out.width="100%", include=F, eval=F}

# Let's attempt an alluvial plot and call it quits:
allSRAFinal %>% count(finalGeography, strictestGeography, finalRace, strictestRace) %>%
ggplot(data = .,
       aes(axis1 = finalGeography, axis2 = strictestGeography, axis3 = finalRace, axis4 = strictestRace, y = n)) +
  scale_x_discrete(limits = c("Original", "Strict", "Strictest"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = finalGeography)) +
  # scale_fill_brewer(palette="Set3") +
    geom_stratum(width=1/3) +
  ylab("Samples") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
  theme(legend.position = "none")
```
