---
title: "Merging the PCA with our calls, part 1"
layout: "post"
tags: "diversity"
output: html_document
date: "2025-05-06"
editor_options: 
  chunk_output_type: console
---

We will finally analyse this data properly. 

```{r setup, include=F}
knitr::opts_knit$set(root.dir="~/OneDrive - St Vincent's Institute/Documents/RNA\ Diversity/", echo=T)
# But we also have to do this manually because R studio is stupid: 
setwd(knitr::opts_knit$get("root.dir"))
library(ggplot2)
library(data.table)
library(plyr)
library(reshape)
library(viridis)
library(rentrez)
library(tidyverse)
library(xml2)
library(R.utils)

options(device = "quartz")
```

So, working with Kasper's data I realised that... they have not cleaned out things like Hela etc, which I have. So we're going to try to that, and then we're going to try to overlap their samples with our own list of primary samples, which is going to require some thinking on my part to remember what we were doing in the first place. 

The first thing we need to do is find equivalent identifiers between our data and theirs. They use DRR something something which does not match our SRA identifiers, but I have faith in my ability to reconcile the two. First we check if the identifiers were in our raw SRA dumps:

Ultimately we want to 
1. clean the data to remove known cell lines like HeLa
2. compare with a sankey between our calls from meta data and his calls from genotype, including unassigned. 
3. can also check how often the same biosample is assigned different ancestries by Kasper lol. 

```{r read-data-and-ids}
fromKasper <- read.csv("from_Kasper_suppTable3.csv")
allSRAFinal <- readRDS("20240901_allSRAFinal_for_plotting.rds") #

head(fromKasper)
head(allSRAFinal)
```

And the sad answer is no... their ID is a run ID, whereas mine is a biosample ID... but they should not be hard to link... if I could work out what fields I want. 

But! We do have SRA.Study ids in common, so we just need to grab the biosample ID that corresponds to Kasper's run ID, and we are good to go for all downstream merging!

```{r convert-run-to-biosample-1, eval=F}
# We start with a test ID from Kasper, and a test ID from me for the same sample:
kasperID <- "ERR188055" # from Kasper
igrID <- "SAMEA1573329" # from me -

kOnion <- entrez_search(db = "sra", term = kasperID)
iOnion <- entrez_search(db = "sra", term = igrID)

kOnion[1]
iOnion[1]
```

So that can be done, but we gotta grab the right IDs only, which makes it harder. Let's start with two or three studies, then. 

It also seems like we can cut down on the number of calls to NCBI by looking at studies instead of each run individually, which should make this faster and less embarrassing.

```{r from-bio-project-to-everything-1}
# Function to get Runs and BioSamples for a BioProject
# Times out after five minutes because sometimes you just gotta know when to give up. Removed the multiple attempts, I'll just make a list of missing BPs later and rerun.
get_runs_and_biosamples <- function(bp_id, timeout=180) {
  result <- tryCatch({
    withTimeout({ 
      # For debugging:
      startTime <- Sys.time()
      print(paste0("Now processing ", bp_id))

      # Let's not get banned from NCBI... they say less than 3 queries per second, so... 
      Sys.sleep(1)

      # Search for BioProject in SRA (with web history)
      search_res <- entrez_search(db = "sra", term = paste0(bp_id, "[BioProject] AND TRANSCRIPTOMIC[Src]"), use_history = TRUE, retmax = 2000)
    
      if (length(search_res$ids) == 0) {
        return(data.frame(Run = NA, BioSample = NA, BioProject = bp_id, stringsAsFactors = FALSE))
      }
    
      # Retrieve webenv and query_key for pagination
      webenv <- search_res$web_history$WebEnv
      query_key <- search_res$web_history$QueryKey
      
      # Initialize variables for fetching results in chunks
      chunk_size <- 200
      all_results <- list()
      batch_start <- 0  # Initial offset for pagination
    
      repeat {
        # Fetch results in chunks using WebEnv and Query Key
        fetch_res <- entrez_fetch(db = "sra", web_history = search_res, rettype = "xml", WebEnv = webenv, query_key = query_key, parsed=FALSE, retstart = batch_start, retmax = chunk_size)
        xml_doc <- read_xml(fetch_res)
    
        # Extract experiment packages (which contains multiple runs)
        packages <- xml_find_all(xml_doc, ".//EXPERIMENT_PACKAGE")
        
        # Extract Run and BioSample info for each package
        records <- lapply(packages, function(pkg) {
          run_nodes <- xml_find_all(pkg, ".//RUN_SET/RUN")
          run_accessions <- xml_attr(run_nodes, "accession")
          
          biosample_node <- xml_find_first(pkg, ".//EXTERNAL_ID[@namespace='BioSample']")
          biosample_id <- if (!is.na(biosample_node)) xml_text(biosample_node) else NA
          
          data.frame(Run = run_accessions, BioSample = biosample_id, BioProject = bp_id, stringsAsFactors = FALSE)
        })
      
        all_results <- append(all_results, records)
        
        # Increment to fetch the next batch
        batch_start <- batch_start + chunk_size
        
        # Stop fetching if fewer results are returned than requested
        if (length(records) < chunk_size) {
          break
        }
      }
    
      # Combine results from all batches
      endTime <- Sys.time()
      print(paste0("Finished processing ", bp_id, ". It took ", round(endTime - startTime, 2), " seconds."))
      return(bind_rows(all_results))
    }, timeout = timeout, onTimeout = "error")
  }, error = function(e) {
    message(paste("Error or timeout for BioProject", bp_id, ":", e$message))
    return(data.frame(Run = NA, BioSample = NA, BioProject = bp_id, stringsAsFactors = FALSE))
  })

  return(result)
}


```

The function has been defined and now let's do a test run with a small number of bioProjects (nb actually not run, just included here for completeness):

```{r from-bioproject-to-everything-2, eval=T}
# For debugging:
kasperBPs <- c("DRP000366", "DRP000425", "DRP000499", "DRP000527", "DRP000622", "DRP000665", "DRP000929", "DRP000987", "DRP001048", "DRP001149", "DRP001150", "DRP001194", "DRP001219", "DRP001220", "DRP001280", "DRP001797", "DRP001919", "DRP001953", "DRP002315", "DRP005280") # Has one of the edge cases
kasperBP <- "DRP000499"
testBPs <- lapply(kasperBP, get_runs_and_biosamples) # Runs fine, do a real one now

testBPs <- lapply(kasperBPs, get_runs_and_biosamples)
testIDs <- bind_rows(testBPs)

#Let's compare with what he had. First, an inner_join
sraToKasper <- inner_join(testIDs, fromKasper, join_by("Run" == "sub_pop"))
dim(sraToKasper) # That gives 711 

dim(fromKasper[fromKasper$pop %in% kasperBPs,]) # and that also gives 711... 

fromKasperSlim <- fromKasper[fromKasper$pop %in% kasperBPs,]

table(testIDs$BioProject)
table(fromKasperSlim$pop)

table(testIDs$BioProject) - table(fromKasperSlim$pop)

kasperComplete <- inner_join(fromKasperSlim, testIDs, join_by("sub_pop" == "Run"))
allSRAFinalKasper <- inner_join(allSRAFinal, kasperComplete, join_by("BioSample" == "BioSample"))

dim(allSRAFinalKasper) 

# But that's too many rows because we're operating at the Run level rather than the BioSample level, so let's clean that mess up:
allSRAFinalKasper <- allSRAFinalKasper %>% distinct(BioSample, .keep_all=TRUE)
dim(allSRAFinalKasper)

# Can't help it I wanna tabulate already:
allSRAFinalKasper %>% count(pred_pop, strictestGeography)
allSRAFinalKasper %>% count(pred_pop, strictestRace)
allSRAFinalKasper %>% filter(pred_pop == "SAS")
```

Ok this works - took some debugging now. So now let's run the entire set of queries, 100 at a time so it doesn't completely burn down:

```{r from-bio-project-to-everything-3, eval=F}

# Batch processing function
process_in_batches <- function(bioProjects, batch_size = 100, delay = 1, file_prefix) {
  # Split the BioProjects list into smaller chunks
  batches <- split(bioProjects, ceiling(seq_along(bioProjects) / batch_size))
  
  # Process each batch sequentially
  all_results <- list()
  
  total_batches <- length(batches)  # Total number of batches
  
  for (i in seq_along(batches)) {
    # Process each BioProject in the batch
    batch <- batches[[i]]
    batch_results <- lapply(batch, get_runs_and_biosamples)
    
    # Combine the results of this batch
    all_results <- append(all_results, batch_results)

    # Save progress after each batch (you could save it as CSV or RDS)
    final_results <- bind_rows(all_results)
    write.csv(final_results, paste0(file_prefix, "bioproject_biosample_runlist.csv"), row.names = FALSE)  # Save as CSV

    # Print a message after each batch is done
    message(paste("Batch", i, "of", total_batches, "completed."))

    # Optional: Add delay after each batch (to avoid rate limits)
    Sys.sleep(delay)  # Delay after each batch
  }
  
  # Combine all the results into one data frame
  final_results <- bind_rows(all_results)
  
  return(final_results)
  }

# Run the batch processing for the BioProjects
# bpsToGrab <- unique(fromKasper$pop)[1:100] # for testing some of the larger ones...
bpsToGrab <- unique(fromKasper$pop)

# If this fails again we can limit the number of queries to those with more than 10 runs...

allBPs <- process_in_batches(bpsToGrab, batch_size = 25, delay = 1, file_prefix = "all_")
```

```{r from-bio-project-to-everything-4, eval=F}
# We read everything in and do a repeat run for everything that failed the first time around (read it in because we are not interested in rerunning the previous bit of code, even by accident)

allBPs <- read.csv("all_bioproject_biosample_runlist.csv", header=T)
dim(allBPs)

fromKasperCheck <- left_join(fromKasper, allBPs, join_by("sub_pop" == "Run")) %>% filter(is.na(BioSample))
dim(fromKasperCheck) # That's a lot of missingness... what BPs go with it?

bpsToRepeat <- fromKasperCheck %>% count(pop) # 61 BPs missing some sort of info

repeatBPs <- process_in_batches(bpsToRepeat$pop, batch_size = 25, delay = 1, file_prefix = "repeats_") # Some of these are clearly doomed and return NA, but others may be salvageable?

processedBPs <- bind_rows(allBPs, repeatBPs)
fromKasperCheck <- left_join(fromKasper, processedBPs, join_by("sub_pop" == "Run")) %>% filter(is.na(BioSample)) 
dim(fromKasperCheck) # That's a lot of missingness... what BPs go with it?

bpsToThreepeat <- fromKasperCheck %>% count(pop) # Down to 48...

# I believe in third chances! There's two huge ones in there that time out, so we can run those two individually with a longer time out and remove them from bpsToThreepeat so not everyone gets 10 mins before timing out:
bigBP <- "SRP200058"
biggerBP <- "SRP166108"

hugeBPs <- c("SRP166108", "SRP200058")

bpsToThreepeat <- bpsToThreepeat[!bpsToThreepeat$pop %in% hugeBPs,]

bigBPout <- get_runs_and_biosamples(bigBP, timeout = 300)
biggerBPout <- get_runs_and_biosamples(biggerBP, timeout = 600)

threepeatBPs <- process_in_batches(bpsToThreepeat$pop, batch_size = 25, delay = 1, file_prefix = "threepeats_")

processedBPs <- bind_rows(processedBPs, bigBPout)
processedBPs <- bind_rows(processedBPs, biggerBPout)
processedBPs <- bind_rows(processedBPs, threepeatBPs)
fromKasperCheck <- left_join(fromKasper, processedBPs, join_by("sub_pop" == "Run")) %>% filter(is.na(BioSample)) 
dim(fromKasperCheck) # From repeating things it really seems like some just fail out of spite, really, so there's no value in re-running things - we rescued most of the missing IDs.

processedBPs <- processedBPs %>% distinct(Run, .keep_all = TRUE)
dim(processedBPs)
write.csv(processedBPs, file="final_bioproject_biosample_runlist.csv", quote=F, row.names=F)

# And just so no one complains, here's what is missing:
fromKasperCheck %>% count(pred_pop)
```

If I was feeling devious I would check:

1. diversity in his data, after cleaning it down to biosample, 10 samples per study, and removing obvious cell lines (HeLa etc)
2. agreement between him and us, as either alluvial or confusion matrix - this can also lead to point 3 below.
3. sample swaps in Run IDs, but that's really just for my own personal edification - doesn't fit with the paper.
