---
title: "Grabbing country for all institutions"
author: "Irene Gallego Romero"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=F}
knitr::opts_knit$set(root.dir="~/Library/CloudStorage/OneDrive-St.Vincent'sInstitute/Documents/RNA Diversity", echo=T)
# But we also have to do this manually because R studio is stupid: 
setwd(knitr::opts_knit$get("root.dir"))
library(data.table)
library(plyr)
library(reshape)
library(tidyverse)
library(stringdist)
library(purrr)
library(rentrez)

options(device = "quartz")
```

Let's try to use ROR to grab country identifiers for all institutions that we can then include in the map. This beats my manual approach from a while ago by, like, a lot... save that it is noisier with the edge cases, because I can resolve them manually and it will not. 

We're going to use the [ROR](https://ror.org/) data dump for this. I initially grabbed version v1.65-2025-05-05-ror-data from Zenodo on 14.05.2025.

Update: In Feb 2026 I switched to the current release, v2.2-2026-01-29-ror-data, which contains about 5k more institutions than the previous one. 

```{r read-big-dataset, eval=T}
allSRAFilt <- readRDS("20250518_allSRACleanRazi_left_join_all_samples.rds")
# And we also sanity check the institution:
table(is.na(allSRAFilt$Center.Name))

allSRAFilt %>% filter(Center.Name == "GEO") %>% summarise(distinct_studies = n_distinct(SRA.Study))
```

Next we need to grab all real submission places for the IDs that match GEO, which is going to be slow - there are 8000 of them - but worthwhile. Bringing the code over from 05, but asking our friend gpt to build a wrapper function around it in the spirit of other GEO queries to save as we go, because 8000 unique studies is a lot of studies. 

Currently not run and set to just test the first 10 IDs for sanity checking.

```{r grab-real-geo-submitters-1, eval=F}
length(unique(allSRAFilt$Center.Name))

siteStudy <- allSRAFilt %>% count(Center.Name, SRA.Study)
geoStudies <- siteStudy[siteStudy$Center.Name %in% "GEO",]$SRA.Study 

process_geo_submitters <- function(studyList, output_prefix = "final_Site_descriptors") {
  
    with_retry <- function(expr, max_tries = 3, wait = 1) {
    for (try in seq_len(max_tries)) {
      res <- tryCatch({
        eval(expr)
      }, error = function(e) {
        if (try == max_tries) stop(e) else NULL
      })
      if (!is.null(res)) return(res)
      Sys.sleep(wait)
    }
    return(NULL)
  }
  
  getBPid <- vector("list", length(studyList))
  names(getBPid) <- studyList
  
for (i in seq_along(studyList)) {
    tryCatch({
      sra_id <- with_retry(quote(entrez_search(db = "sra", term = studyList[i])$ids[1]))
      if (!is.null(sra_id)) {
        getBPid[[i]] <- with_retry(quote(entrez_link(dbfrom = "sra", id = sra_id, db = "all")$links$sra_bioproject_all))
      }
    }, error = function(e) {
      message(sprintf("Error at index %d for study %s: %s", i, studyList[i], e$message))
      getBPid[[i]] <- NA
    })
    
    
    if (i %% 100 == 0) {
      saveRDS(getBPid, file = sprintf("tmp_%s_getBPid_%03d.rds", output_prefix, i))
      message(sprintf("Saved getBPid progress at %d / %d", i, length(studyList)))
    }
  }
  
  getSubmitter <- vector("character", length(getBPid))
  names(getSubmitter) <- names(getBPid)
  
  for (i in seq_along(getBPid)) {
    tryCatch({
      getSubmitter[i] <- with_retry(quote(
        extract_from_esummary(entrez_summary(db = "bioproject", id = getBPid[[i]][[1]][1]), "submitter_organization")
      ))
    }, error = function(e) {
      message(sprintf("Error extracting submitter at index %d: %s", i, e$message))
      getSubmitter[i] <- NA
    })
    
    if (i %% 100 == 0) {
      saveRDS(getSubmitter, file = sprintf("tmp_%s_getSubmitter_%03d.rds", output_prefix, i))
      message(sprintf("Saved getSubmitter progress at %d / %d", i, length(getBPid)))
    }
  }
  
  getSubmitterdf <- data.frame(
    SRA = names(getSubmitter),
    submitter = unname(getSubmitter),
    stringsAsFactors = FALSE
  )
  
  outfile <- sprintf("%s_submitter_lookup.tsv", output_prefix)
  write.table(getSubmitterdf, file = outfile, quote = FALSE, row.names = FALSE, sep = "\t")
  message("Submitter lookup table written to: ", outfile)
  
  return(getSubmitterdf)
}

# submitter_info <- process_geo_submitters(geoStudies[1:10], output_prefix = "20250603_final_Site_descriptors")
```

```{r grab-real-geo-submitters-2}
submitter_info <- read.table("20250603_final_Site_descriptors_submitter_lookup.tsv", sep="\t", header=T, stringsAsFactors = F, quote = "")

# The final 6 rows are not associated with SRA IDs for some reason, so we drop them:
submitter_info <- submitter_info[1:(nrow(submitter_info)-6),]

# And now we update
allSRAFilt$finalSite <- NA

for (i in 1:nrow(submitter_info)){
  allSRAFilt[allSRAFilt$SRA.Study %in% submitter_info$SRA[i],]$finalSite <- submitter_info$submitter[i]    
}

allSRAFilt$finalSite <- coalesce(allSRAFilt$finalSite, allSRAFilt$Center.Name)

saveRDS(allSRAFilt, file="20260209_allSRAFilt_Razi_left_institution_intermediate.rds")

# Is GEO really gone???
length(unique(allSRAFilt$Center.Name))
length(unique(allSRAFilt$finalSite)) # Well that definitely increased, so that's good. 

allSRAFilt %>% count(finalSite, sort=T) %>% head(n = 50)
allSRAFilt %>% count(Center.Name, sort=T) %>% head(n = 50)

institutions <- data.frame(name = tolower(unique(allSRAFilt$finalSite)))
dim(institutions)
```

Now we start working with the ROR data, which will be new and exciting... 

```{r read-and-parse-ror, eval=T}
# rorDataOld <- read.csv("ROR_data/v1.65-2025-05-05-ror-data_schema_v2.csv", stringsAsFactors = F)
rorData <- read.csv("ROR_data/v2.2-2026-01-29-ror-data.csv", stringsAsFactors = F)

# Extract useful fields: name, aliases, acronyms, country
rorClean <- data.frame(
  label = tolower(rorData$names.types.ror_display),
  acronyms = tolower(rorData$names.types.acronym),
  aliases = tolower(rorData$names.types.alias),
  country = rorData$locations.geonames_details.country_name,
  stringsAsFactors = FALSE
)

# Replace NAs with empty strings for merging fields in the next few steps...
rorClean[is.na(rorClean)] <- ""

#Build unified name list
rorClean$all_names <- paste(rorClean$label, rorClean$aliases, rorClean$acronyms, sep = ";")
ror_lookup <- rorClean %>%
  rowwise() %>%
  mutate(name_list = strsplit(all_names, ";")) %>%
  unnest(name_list) %>%
  filter(name_list != "") %>%
  select(name_list, country) %>%
  distinct()

# And now let's clean it up... I suspect the opening "the" is messing up a lot of matches, so let's get rid of it and ask ourselves, how much worse can that make things, plus all the language code info that will defs ruin the matching:
ror_lookup <- ror_lookup %>%
  mutate(name_list = gsub("^the ", "", name_list)) %>% 
  mutate(name_list = gsub("no_lang_code: ", "", name_list)) %>%
  mutate(name_list = gsub("[a-z][a-z]: ", "", name_list)) 

# Keeping only the element after the final comma, in case there are multiple, since I often see the institution that could be matched is there - this should help with country matching but needs to be handled carefully when trying to unite the columns later... or does it?
institutions <- institutions %>%
  mutate(name = gsub("^the ", "", name)) %>% 
  mutate(lastName = map_chr(str_split(name, ","), ~ tail(.x, 1))) %>% 
  mutate(lastName = gsub("^the ", "", lastName)) %>%
  mutate(lastName = gsub("^ ", "", lastName))

# Match function using Jaro-Winkler distance (measures edit distance, 0 is identical whereas 1 is totally different, favours matches from the beginning, commonly used in fuzzy text matching)
find_best_match <- function(inst_name, ror_names) {
  if (is.na(inst_name) || inst_name == "" || length(ror_names) == 0) {
    return(c(match_name = NA, distance = NA))
  }

  # Compute distances
  distances <- stringdist(inst_name, ror_names, method = "jw")

  # If all distances are NA or vector is empty
  if (all(is.na(distances))) {
    return(c(match_name = NA, distance = NA))
  }

  # Get best match safely
  best_idx <- which.min(distances)
  return(c(match_name = ror_names[best_idx], distance = distances[best_idx]))
}

matches <- t(sapply(institutions$name, find_best_match, ror_lookup$name_list))

lastMatches <- t(sapply(institutions$lastName, find_best_match, ror_lookup$name_list))

matchedNames <- institutions %>%
  bind_cols(matches) %>%
  left_join(ror_lookup, by = c("match_name" = "name_list"))

matchedLastNames <- institutions %>%
  bind_cols(lastMatches) %>%
  left_join(ror_lookup, by = c("match_name" = "name_list"))

allMatchedNames <- left_join(matchedNames, matchedLastNames, by = c("name" = "name"), suffix = c(
  "_name", "_lastName"))

allMatchedNames <- allMatchedNames[,c(1,2,3,4,5,7,8,9)]
nrow(allMatchedNames)

allMatchedNames %>% distinct(name, lastName_name, country_name, country_lastName, .keep_all = T) %>% head(n = 20)

write.table(allMatchedNames, file="20260209_institution_country_matches.tsv", quote=F, row.names=F, sep="\t")
```

Now let's take a look at the output and see if we can make any sense of it... There are a lot of of dumpster fires but many are easy to fix, though how do I justify bothering? How reliable is this overall?

I first defined allMatchedNamesClean as being <= 0.1, but that was still extremely noisy, so I am now sticking with 0.05.

```{r explore-name-matches-1, eval=T}
allMatchedNames <- read.table("20260209_institution_country_matches.tsv", header=T, quote="", sep="\t", comment.char = "")

# The internet won't tell me what's a 'good' match so let's just eyeball some different subsets
allMatchedNames %>% filter(distance_name > 0) %>% arrange(distance_name) %>% head(n = 50)
allMatchedNames %>% filter(distance_name > 0.05) %>% arrange(distance_name) %>% head(n = 50)
allMatchedNames %>% filter(distance_name > 0.1) %>% arrange(distance_name) %>% head(n = 50)
allMatchedNames %>% filter(distance_name > 0.2) %>% arrange(distance_name) %>% head(n = 50)
allMatchedNames %>% filter(distance_name > 0.3) %>% arrange(distance_name) %>% head(n = 50)

allMatchedNames %>% filter(distance_name == 0) %>% nrow()

allMatchedNamesClean <- allMatchedNames %>% filter(distance_name <= 0.05 | distance_lastName <= 0.05)

allMatchedNamesClean %>% filter(distance_name > 0) %>% arrange(distance_name) %>% head(n = 50)
allMatchedNamesClean %>% filter(distance_name > 0.05) %>% arrange(distance_name) %>% head(n = 50)
allMatchedNamesClean %>% filter(distance_name > 0.1) %>% arrange(distance_name) %>% head(n = 50)
allMatchedNamesClean %>% filter(distance_name > 0.2) %>% arrange(distance_name) %>% head(n = 50)
allMatchedNamesClean %>% filter(distance_name > 0.3) %>% arrange(distance_name) %>% head(n = 50)

allMatchedNamesClean %>% filter(distance_name == 0) %>% nrow()

allMatchedNamesClean %>% filter(distance_lastName > 0) %>% arrange(distance_lastName) %>% head(n = 50)
allMatchedNamesClean %>% filter(distance_lastName > 0.05) %>% arrange(distance_lastName) %>% head(n = 50)
allMatchedNamesClean %>% filter(distance_lastName > 0.1) %>% arrange(distance_lastName) %>% head(n = 50)

allMatchedNamesClean %>% filter(distance_lastName == 0) %>% nrow() # Clearly much better
allMatchedNames %>% filter(distance_name == 0) %>% nrow()

allMatchedNamesClean$better <- ifelse(allMatchedNamesClean$distance_lastName < allMatchedNamesClean$distance_name, "lastName", "name")
table(allMatchedNamesClean$better)

allMatchedNamesClean$countryMatch <- allMatchedNamesClean$country_lastName ==  allMatchedNamesClean$country_name
table(allMatchedNamesClean$countryMatch) # YIKES

write.table(allMatchedNamesClean, file="20260209_institution_country_matches_0.05.tsv", quote=F, row.names=F, sep="\t")
```

There's some pretty outrageous mistakes that would require manual clean up, and some really perplexing ones, eg university of maryland going to university of malakand (tho if you look at the raw data there isn't a single sensible entry for maryland...). Garvan also failing very hard. I think we have to apply a clear cutoff of some sort and just be like, well, that sucks, but that's the overall issue with the metadata too. 

```{r more-strigent-distances}
# Let's consider a JW distance of 0 only... doesn't seem to make much of a difference compared to the numbers above, shockingly. 
allMatchedNamesCleanest <- allMatchedNames %>% filter(distance_name <= 0.00 | distance_lastName <= 0.00)

allMatchedNamesCleanest$better <- ifelse(allMatchedNamesCleanest$distance_lastName < allMatchedNamesCleanest$distance_name, "lastName", "name")
table(allMatchedNamesCleanest$better)

allMatchedNamesCleanest$countryMatch <- allMatchedNamesCleanest$country_lastName ==  allMatchedNamesCleanest$country_name
table(allMatchedNamesCleanest$countryMatch)
```


```{r explore-name-matches-2}
dim(allMatchedNamesCleanest)
dim(allMatchedNamesClean)
dim(institutions)

length(unique(allMatchedNamesClean$name))
length(unique(allMatchedNamesClean$match_name_name))
#Uh.... some things are clearly missing here and we don't like that. But what is missing??? Before we delve into that, let's quickly look at some of the matches across confidence bins:
length(unique(allMatchedNamesCleanest$name))
length(unique(allMatchedNamesCleanest$match_name_name))

allMatchedNamesClean$distance_name <- as.numeric(allMatchedNamesClean$distance_name)
allMatchedNamesClean$distance_lastName <- as.numeric(allMatchedNamesClean$distance_lastName)

summary(allMatchedNamesClean$distance_name)
summary(allMatchedNamesClean$distance_lastName)

allMatchedNamesClean %>% arrange(desc(distance_lastName)) %>% filter(better != "name") %>% head(n = 50)

length(unique(allMatchedNamesClean$country_name))
length(unique(allMatchedNamesClean$country_lastName))
intersect(allMatchedNamesClean$country_name, allMatchedNamesClean$country_lastName)

length(unique(allMatchedNamesCleanest$country_name))
length(unique(allMatchedNamesCleanest$country_lastName))
intersect(allMatchedNamesCleanest$country_name, allMatchedNamesCleanest$country_lastName)

allMatchedNamesClean %>% count(country_name, sort=T) %>% head(n=20) 
allMatchedNamesClean %>% count(country_lastName, sort=T) %>% head(n=20)

allMatchedNamesCleanest %>% count(country_name, sort=T) %>% head(n=20) 
allMatchedNamesCleanest %>% count(country_lastName, sort=T) %>% head(n=20)

allMatchedNamesClean %>% count(country_name, sort=T) %>% tail(n=20) 
allMatchedNamesClean %>% count(country_lastName, sort=T) %>% tail(n=20)

allMatchedNamesCleanest %>% count(country_name, sort=T) %>% tail(n=20) 
allMatchedNamesCleanest %>% count(country_lastName, sort=T) %>% tail(n=20)

# Let's look at some of the more unusual countries: 
allMatchedNamesClean[allMatchedNamesClean$country_name %in% "Bolivia",] # Correctly parsed as USA by lastName
allMatchedNamesClean[allMatchedNamesClean$country_lastName %in% "Rwanda",] # Correct match on lastName
allMatchedNamesClean[allMatchedNamesClean$country_lastName %in% "Aruba",] # Looks wrong
allMatchedNamesClean[allMatchedNamesClean$country_lastName %in% "Costa Rica",] # Should probably be Mexico (UNAM)
allMatchedNamesClean[allMatchedNamesClean$country_lastName %in% "United Arab Emirates",] # All three correct... 
```

Listen... we do our best. I think it's as good as it is going to get unless we wanted to manually tweak too many things. So now we just need to merge this back into allSRAFilt and go from there... 

There's still the problem of deciding what to do with institutions that are present in multiple countries - which I think we discard, even if it seems like madness because we have strong priors!

```{r merge-back-to-allSRA}
# And now we update. The filtering means these are in complete disarray, so we need to clean up a bit first... Institutions was parsed a tiny bit, so we create a similar vector in allSRA and attempt to match by that, starting from finalSite, which has dealt with the GEO mess

# allSRAFilt <- readRDS(file="20260209_allSRAFilt_Razi_left_institution_intermediate.rds")

allSRAFilt <- allSRAFilt %>%
  mutate(tempInst = gsub("^the ", "", tolower(finalSite))) 

betterLast <- allMatchedNamesClean %>% filter(better == "lastName")
betterName <- allMatchedNamesClean %>% filter(better == "name")

betterLast %>% group_by(match_name_lastName) %>% summarise(countries = n_distinct(country_lastName)) %>% summarise(mean(countries), max(countries)) 
betterName %>% group_by(match_name_name) %>% summarise(countries = n_distinct(country_name))%>% summarise(mean(countries), max(countries)) 

strictLast <- allMatchedNamesCleanest %>% filter(better == "lastName")
strictName <- allMatchedNamesCleanest %>% filter(better == "name")

strictLast %>% group_by(match_name_lastName) %>% summarise(countries = n_distinct(country_lastName)) %>% summarise(mean(countries), max(countries)) 
strictName %>% group_by(match_name_name) %>% summarise(countries = n_distinct(country_name))%>% summarise(mean(countries), max(countries)) 

# Matches with only one country
singleLast <- betterLast %>% group_by(match_name_lastName) %>% summarise(countries = n_distinct(country_lastName)) %>% filter(countries == 1)
singleName <- betterName %>% group_by(match_name_name) %>% summarise(countries = n_distinct(country_name)) %>% filter(countries == 1)

singleLastStrict <- strictLast %>% group_by(match_name_lastName) %>% summarise(countries = n_distinct(country_lastName)) %>% filter(countries == 1)
singleNameStrict <- strictName %>% group_by(match_name_name) %>% summarise(countries = n_distinct(country_name)) %>% filter(countries == 1)

betterLast <- betterLast[betterLast$match_name_lastName %in% singleLast$match_name_lastName,]
betterName <- betterName[betterName$match_name_name %in% singleName$match_name_name,]

strictLast <- strictLast[strictLast$match_name_lastName %in% singleLastStrict$match_name_lastName,]
strictName <- strictName[strictName$match_name_name %in% singleNameStrict$match_name_name,]

# And now the scary merge
allSRAFilt$lastCountry <- NA
allSRAFilt$nameCountry <- NA

allSRAFilt$lastCountryStrict <- NA
allSRAFilt$nameCountryStrict <- NA

for (i in 1:nrow(betterLast)){
  allSRAFilt[allSRAFilt$tempInst %in% betterLast$name[i],]$lastCountry <- betterLast$country_lastName[i]
}

for (i in 1:nrow(betterName)){
  allSRAFilt[allSRAFilt$tempInst %in% betterName$name[i],]$nameCountry <- betterName$country_name[i]
}

for (i in 1:nrow(strictLast)){
  allSRAFilt[allSRAFilt$tempInst %in% strictLast$name[i],]$lastCountryStrict <- strictLast$country_lastName[i]    
}

for (i in 1:nrow(strictName)){
  allSRAFilt[allSRAFilt$tempInst %in% strictName$name[i],]$nameCountryStrict <- strictName$country_name[i]    
}

table(allSRAFilt$lastCountry)
table(allSRAFilt$nameCountry)

table(allSRAFilt$lastCountryStrict)
table(allSRAFilt$nameCountryStrict)

table(is.na(allSRAFilt$lastCountry), is.na(allSRAFilt$nameCountry)) # How many double assignments...None! What a shock!
table(is.na(allSRAFilt$lastCountryStrict), is.na(allSRAFilt$nameCountryStrict)) # And with this one?

allSRAFilt$finalCountry <- coalesce(allSRAFilt$nameCountry, allSRAFilt$lastCountry)
allSRAFilt$strictCountry <- coalesce(allSRAFilt$nameCountryStrict, allSRAFilt$lastCountryStrict)

table(allSRAFilt$finalCountry == allSRAFilt$strictCountry) # Not sure why this doesn't pick the NAs...

allSRAFilt %>% filter(finalCountry != strictCountry) %>% head() # Makes sense that all of these align perfectly, because after all strictCountry is a subset of finalCountry... 
# allSRAFilt %>% filter(is.na(strictCountry) & !is.na(finalCountry)) %>% head(n = 50)
allSRAFilt %>% filter(is.na(strictCountry) & !is.na(finalCountry)) %>% nrow() # We lose about 30,000 assignments

strictCounts <- allSRAFilt %>% count(strictCountry)
finalCounts <- allSRAFilt %>% count(finalCountry)
allCountries <- full_join(finalCounts, strictCounts, by = c("finalCountry" = "strictCountry"))
names(allCountries) <- c("Country", "nFinal", "nStrict")
allCountries <- allCountries %>%
  mutate(nStrict = ifelse(is.na(nStrict), 0, nStrict),
         diff = nFinal - nStrict,
         fractDiff = diff/nFinal)

allCountries %>% arrange(desc(diff)) %>% head(n = 10)
allCountries %>% arrange(desc(fractDiff)) %>% head(n = 20)

# Quickly checking some of the countries with the biggest differences fractionally
allSRAFilt %>% filter(finalCountry == "Algeria") %>% select(c(1,2,3,56:ncol(allSRAFilt))) # Clear OG fail
allSRAFilt %>% filter(finalCountry == "Tunisia") %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # Looks like a fail
allSRAFilt %>% filter(finalCountry == "Cuba" & is.na(strictCountry)) %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # Super wrong, should be Spain
allSRAFilt %>% filter(finalCountry == "Ecuador" & is.na(strictCountry)) %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # Wrong
allSRAFilt %>% filter(finalCountry == "Belarus" & is.na(strictCountry)) %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # All wrong... 

# And countries with some assignments before and after?
allSRAFilt %>% filter(finalCountry == "Portugal" & is.na(strictCountry)) %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # Very Portuguese institutions
allSRAFilt %>% filter(finalCountry == "Poland" & is.na(strictCountry)) %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # A Polish institutions
allSRAFilt %>% filter(finalCountry == "South Africa" & is.na(strictCountry)) %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # Correct assignment to SA
allSRAFilt %>% filter(finalCountry == "Iran" & is.na(strictCountry)) %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # China??
allSRAFilt %>% filter(finalCountry == "Luxembourg" & is.na(strictCountry)) %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # Should be kept...
allSRAFilt %>% filter(finalCountry == "Czechia" & is.na(strictCountry)) %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # First few are Polish, but many are in Czechia
allSRAFilt %>% filter(finalCountry == "Australia" & is.na(strictCountry)) %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # Definitely Australian...


allSRAFilt %>% filter(strictCountry == "Türkiye") %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # 
allSRAFilt %>% filter(strictCountry == "Iran") %>% select(c(1,2,3,56:ncol(allSRAFilt))) # # 
allSRAFilt %>% filter(strictCountry == "Portugal") %>% select(c(1,2,3,56:ncol(allSRAFilt))) # The assigments to IT are incorrect, but all the other ones are fine
allSRAFilt %>% filter(strictCountry == "Poland") %>% select(c(1,2,3,56:ncol(allSRAFilt))) # Two different institutes of human genetics, one of which seems to be in Germany? Hard to tell, since there's a Polish one and SRA doesn't have much info on the submitters.They also lose a LOT of data compared to 0.05... but a lot of it was being miscoded as Czechia anyhow (all the PAS samples)

# O.0 is very conservative, and penalises things like apostrophes or accents, I bet, which are more likely to be reported in ROR than in SRA. But...


head(allSRAFilt)
allSRAFilt %>% count(finalCountry, sort=T) %>% head(n=20) 
allSRAFilt %>% count(strictCountry, sort=T) %>% head(n=20) 

allSRAFilt <- allSRAFilt %>% select(-lastCountry, -nameCountry, -lastCountryStrict, -nameCountryStrict, -tempInst)

saveRDS(allSRAFilt, file="20260209_allSRAFilt_Razi_left_country.rds")
```


