---
title: "Merging the PCA with our calls"
layout: "post"
tags: "diversity"
output: html_document
date: "2025-05-06"
editor_options: 
  chunk_output_type: console
---

We will finally analyse this data properly. 

```{r setup, include=F}
knitr::opts_knit$set(root.dir="~/OneDrive - St Vincent's Institute/Documents/RNA\ Diversity/", echo=T)
# But we also have to do this manually because R studio is stupid: 
setwd(knitr::opts_knit$get("root.dir"))
library(ggplot2)
library(data.table)
library(plyr)
library(reshape)
library(viridis)
library(rentrez)
library(tidyverse)

options(device = "quartz")
```

So, working with Kasper's data I realised that... they have not cleaned out things like Hela etc, which I have. So we're going to try to that, and then we're going to try to overlap their samples with our own list of primary samples, which is going to require some thinking on my part to remember what we were doing in the first place. 

The first thing we need to do is find equivalent identifiers between our data and theirs. They use DRR something something which does not match our SRA identifiers, but I have faith in my ability to reconcile the two. First we check if the identifiers were in our raw SRA dumps:


Ultimately we want to 
1. clean the data to remove known cell lines like HeLa
2. compare with a sankey between our calls from meta data and his calls from genotype, including unassigned. 
3. can also check how often the same biosample is assigned different ancestries by Kasper lol. 

```{r read-data-and-ids}
fromKasper <- read.csv("from_Kasper_suppTable3.csv")
sraPre <- read.csv('SRA_metadata/SRA_batches/SraRunTable_00.txt',header=T)
allSRAFinal <- readRDS("20240901_allSRAFinal_for_plotting.rds") #

head(fromKasper)
head(sraPre)
head(allSRAFinal)
```

And the sad answer is no... their ID is a run ID, whereas mine is a biosample ID... but they should not be hard to link... if I could work out what fields I want. 

But! We do have SRA.Study ids in common, so we just need to grab the biosample ID that corresponds to Kasper's run ID, and we are good to go!

Also they're in the same order... not sure if that helps us somehow. 

```{r convert-run-to-biosample-1}
# We start with a test ID from Kasper, and a test ID from me for the same sample:
kasperID <- "ERR188055" # from Kasper
igrID <- "SAMEA1573329" # from me -

kOnion <- entrez_search(db = "sra", term = kasperID)
iOnion <- entrez_search(db = "sra", term = igrID)

kOnion[1]
iOnion[1]
```

So that can be done, but we gotta grab the right IDs only, which makes it harder. Let's start with two or three studies, then. 

```{r convert-run-to-biosample-2}
# A single common study ID
commonID <- "SRP001371"

onion <- entrez_search(db = "sra", term = paste0(commonID, "[GPRJ] AND TRANSCRIPTOMIC[SRC]"))

onion <- entrez_summary(db = "sra", commonID)



iOnion <- entrez_search(db = "sra", term = igrID)

kOnion[1]
iOnion[1]


getBPid <- sapply(geoStudies, function(x) entrez_link(dbfrom="sra", id=entrez_search(db = "sra", term = x)$ids[1], db="all")$links$sra_bioproject_all)
getBPid

getSubmitter <- sapply(getBPid, function(x) extract_from_esummary(entrez_summary(db="bioproject", x[[1]][1]), "submitter_organization"))



# I actually don't care about the unique identifiers anymore... this is good enough then?  Let's test with a few more samples that are again matched:
igrIDs <- allSRAFinal$BioSample[1:20] # Looked these up on SRA and grabbed corresponding run IDs manually from run selector, these should match, but we run into a problem that there's a lot more unique run IDs than BioSample ids! Still, should be able to get to 20 in the end!
kasperIDs <- c("SRR097618", "SRR643708", "SRR643709", "SRR643710", "SRR643711", "SRR643712", "SRR643766", "SRR980415", "SRR980416", "SRR980445", "SRR980446", "SRR787508", "SRR787509", "SRR787517", "SRR787518", "SRR787519", "SRR787520", "SRR787521", "SRR787510", "SRR787511", "SRR787512", "SRR097617", "SRR644512", "SRR616719", "SRR620433", "SRR620434", "SRR620435", "SRR620436", "SRR620437", "SRR620438", "SRR1385239", "SRR1385240", "SRR1385241", "SRR1385242", "SRR1385243", "SRR1385244", "SRR1385245", "SRR1385246", "SRR1385247", "SRR1385248", "SRR1385249", "SRR1385250", "SRR6176944", "SRR6176947", "SRR6176950", "SRR6176952", "SRR6176953", "SRR1067126", "SRR1067226", "SRR1067390", "SRR1068371", "SRR1068374", "SRR1068377", "SRR1068379", "SRR1068407", "SRR1143698", "SRR1143699", "SRR1143701", "SRR1143703", "SRR643702", "SRR643703", "SRR643704", "SRR643705", "SRR643706", "SRR643707", "SRR066170", "SRR066171", "SRR066172", "SRR066173", "SRR453391", "SRR453392", "SRR453393", "SRR453394", "SRR453395", "SRR453396", "SRR453397", "SRR453398", "SRR453399", "SRR453400", "SRR453401", "SRR453402", "SRR453403", "SRR453404", "SRR453405", "SRR453406", "SRR453407")

# Let's make sure these are all here, or at least a decent fraction is:
length(kasperIDs)
dim(fromKasper[fromKasper$sub_pop %in% kasperIDs,]) 
```


# It also seems like we can cut down on the number of tests by looking at studies instead of 

iOnion <- sapply(igrIDs, function(x) entrez_search(db = "sra", term = x)$ids)
kOnion <- sapply(kasperIDs, function(x) entrez_search(db = "sra", term = x)$ids)

iOnion2 <- as.data.frame(iOnion)
iOnion2$BioSample <- rownames(iOnion2)

kOnion2 <- as.data.frame(kOnion)
kOnion2$Run <- rownames(kOnion2)

mergedIds <- left_join(iOnion2, kOnion2, join_by("iOnion" == "kOnion"))
mergedIds <- left_join(mergedIds, fromKasper, join_by("Run" == "sub_pop"))

head(mergedIds)

head(fromKasper)

```

# I just need to annotate Kasper's data, not mine, and can then merge that with my own file, both early on for allSRA, to weed out primary cell lines etc, and then for allSRA final to make some more plots...

```{r leftovers}
# This took forever to fix but...
getBPid <- sapply(geoStudies, function(x) entrez_link(dbfrom="sra", id=entrez_search(db = "sra", term = x)$ids[1], db="all")$links$sra_bioproject_all)
getBPid

getSubmitter <- sapply(getBPid, function(x) extract_from_esummary(entrez_summary(db="bioproject", x[[1]][1]), "submitter_organization"))
getSubmitter

getSubmitterdf <- as.data.frame(getSubmitter)
getSubmitterdf$SRA <- rownames(getSubmitterdf)

# And now we add that info back to the relevant column, and replace the GEO calls. I honestly can't remember if I need to vectorise or not but whatever. 
allSRAFinal$finalSite <- NA

for (i in 1:nrow(getSubmitterdf)){
  allSRAFinal[allSRAFinal$SRA.Study %in% getSubmitterdf$SRA[i],]$finalSite <- getSubmitterdf$getSubmitter[i]    
}

allSRAFinal$finalSite <- coalesce(allSRAFinal$finalSite, allSRAFinal$Center.Name)

# Is GEO really gone???
length(unique(allSRAFinal$Center.Name))
length(unique(allSRAFinal$finalSite)) # Well that definitely increased, so that's good. 

siteTable <- allSRAFinal %>% count(finalSite, SRA.Study)

# I still need to work out what is in what country, but at least I have a single table... 
write.table(siteTable, file="final_Site_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")

# Anyhow, let's make sure we only do this once:
saveRDS(allSRAFinal, "allSRAFinalTissuesCenters.rds")
```