---
title: "Diversity project new base file"
layout: "post"
tags: "diversity"
output: html_document
editor_options: 
  chunk_output_type: console
---

This was getting exhausting, so we're moving it to a repo now, so it can be properly versioned and edited and stuff, I got tired of all the duplicate files and the such. The lab notebook will simply contain notes...



This is mostly Amanda's code, except that I noticed when trying to do tissue analyses that a lot of studies use the 'source_name' column to store tissue information, which is annoying but worth rescuing. Rather than trying to work out how to query the API programatically, which didn't work so well, I am going to use her code to recreate the big master file, and then run it through the clean up process again with a couple of extra columns, and go from there. 

```{r old-sra}
knitr::opts_knit$set(root.dir="~/OneDrive - St Vincent's Institute/Documents/RNA\ Diversity/SRA_metadata/", echo=T)
# But we also have to do this manually because R studio is stupid: 
setwd(knitr::opts_knit$get("root.dir"))
library(plyr)

#########
# combine data extraction batches - pre 2019
#########

files1<-paste0("0", seq(1:9))
files2<-10:15
files<-c(files1,files2)

data_sub=read.csv('SRA_batches/SraRunTable_00.txt',header=T)

for (i in 1:length(files)){
  tmp_sub=read.csv(paste('SRA_batches/SraRunTable_',files[i],'.txt',sep=''),header=T)
  data_sub=rbind.fill(data_sub,tmp_sub)
  print(i)
}

missing<-as.data.frame(apply(data_sub,2,function(x) length(which(x=='' | is.na(x)))))
missing$prop<-missing[,1]/dim(data_sub)[1]
hist(missing$prop)
write.table(as.data.frame((missing)),'missing_metadata_20240613.txt',quote=F,sep='\t')

data_sub2=data_sub[,c('SRA.Study','Center.Name','BioSample','Organism','LibrarySource','Cell_type','Tissue','Cell_line','disease','ETHNICITY','disease_state','Diagnosis','tissue_type','clinical_information','RACE','Site','health_state','Population','tissue_source','geo_loc_name_country','geo_loc_name_country_continent','DONOR_HEALTH_STATUS','DONOR_ETHNICITY','geo_loc_name','reported_race','source_tissue',"ReleaseDate","create_date","ancestry","donor_id", "biosample_type", "COUNTRY", "tissue.cell_type_source", "source_material_cell_type", "source_material", "source_name", "Sample_type", "Sampling_site")]

write.table(data_sub2,'8Sep23_run_selector_metadata.txt',quote=F,row.names=F,sep='\t')

```

I grabbed a bunch of new columns that may or may not be promising:
"biosample_type", "COUNTRY", "tissue.cell_type_source", "source_material_cell_type", "source_material", "source_name", "Sample_type", "Sampling_site"

```{r new-sra}

#########
# combine data extraction batches - post 2019
#########
files<-seq(2:21)

data_sub=read.csv('SRA_post2019/SraRunTable_1.txt',header=T)

for (i in 1:length(files)){
  tmp_sub=read.csv(paste('SRA_post2019/SraRunTable_',files[i],'.txt',sep=''),header=T)
  data_sub=rbind.fill(data_sub,tmp_sub)
  print(i)
}

data_sub2=data_sub[,c('SRA.Study','Center.Name','BioSample','Organism','LibrarySource','Cell_type','Tissue','Cell_Line','disease','ETHNICITY','disease_state','Diagnosis','tissue_type','clinical_information','RACE','Site','health_state','Population','tissue_source','geo_loc_name_country','geo_loc_name_country_continent','geo_loc_name','reported_race','source_tissue',"ReleaseDate","create_date","ancestry","race.ethnicity","primary_race",'geographical_location',"donor_id")]

data_sub2=data_sub[,c('SRA.Study','Center.Name','BioSample','Organism','LibrarySource','Cell_type','Tissue','Cell_line','disease','ETHNICITY','disease_state','Diagnosis','tissue_type','clinical_information','RACE','Site','health_state','Population','tissue_source','geo_loc_name_country','geo_loc_name_country_continent','DONOR_HEALTH_STATUS','DONOR_ETHNICITY','geo_loc_name','reported_race','source_tissue',"ReleaseDate","create_date","ancestry","donor_id", "biosample_type", "COUNTRY", "tissue.cell_type_source", "source_material_cell_type", "source_material", "source_name", "Sample_type", "Sampling_site")]

write.table(data_sub2,'8Sep23_run_selector_metadata_post2019.txt',quote=F,row.names=F,sep='\t')

#########
# combine ALL samples from pre and post 2019
#########

data_subA=read.delim('8Sep23_run_selector_metadata.txt')
data_subB=read.delim('8Sep23_run_selector_metadata_post2019.txt')
names(data_subB)[8]<-'Cell_line'
data_sub2=rbind.fill(data_subA,data_subB)
data_sub2=subset(data_sub2,Organism=='Homo sapiens')
data_sub2=unique(data_sub2)

#########
# filtering
#########

# filter for studies with 10 or more samples, as we did for manual review
data_sub2_size<-as.data.frame(table(data_sub2$SRA.Study))
data_sub2_size<-subset(data_sub2_size,Freq>9)
data_sub2=subset(data_sub2,SRA.Study %in% data_sub2_size$Var1)

# reformat year
data_sub2$ReleaseDate2<-as.POSIXct(data_sub2$ReleaseDate)
data_sub2$ReleaseYear<-format(data_sub2$ReleaseDate2, "%Y")
tmp<-subset(byyear,year<2023)

# remove cell lines to get just primary samples for some analyses
data_sub3<-subset(data_sub2,Cell_line=='' | is.na(Cell_line) | Cell_line=='<NA>' | Cell_line=='NA')
cell_line<-as.data.frame(t(apply(data_sub3[,c('BioSample','Cell_type','Tissue')],1,function(u) grepl("cell line",u))))
cell_line$sum<-apply(cell_line,1,function(x) length(which(x=='TRUE')))
cell_line$id<-1:dim(cell_line)[1]
data_sub3<-data_sub3[-c(subset(cell_line,sum>0)$id),]


```