---
title: "Fetching missing sequencing sites."
layout: "post"
tags: "diversity"
output: html_document
editor_options: 
  chunk_output_type: console
---

We wanted to work out who is sequencing what, and for that we need to fill in a lot of the studies that were submitted by GEO to SRA instead of directly by the submitter. This requires a couple of steps:

First we convert from SRA.Study to Bioproject ID. Then we use the Bioproject ID to fetch the summary using NCBI's eutils' package, so we gotta mix up bash and R in here... yeugh. 

```{r setup, include=F}
knitr::opts_knit$set(root.dir="~/OneDrive - St Vincent's Institute/Documents/RNA\ Diversity/", echo=T)
# But we also have to do this manually because R studio is stupid: 
setwd(knitr::opts_knit$get("root.dir"))
library(ggplot2)
library(data.table)
library(plyr)
library(reshape)
library(viridis)
library(rentrez)
library(tidyverse)

options(device = "quartz")
```

Grab all the SRA IDs that were deposited by GEO and let's figure out where we go from there...

```{r depositing-site, include=F, eval=F}
allSRAFinal <- readRDS("allSRAFinalTissues.rds")
length(unique(allSRAFinal$Center.Name))

siteStudy <- allSRAFinal %>% count(Center.Name, SRA.Study)
geoStudies <- siteStudy[siteStudy$Center.Name %in% "GEO",]$SRA.Study 

# This took forever to fix but...
getBPid <- sapply(geoStudies, function(x) entrez_link(dbfrom="sra", id=entrez_search(db = "sra", term = x)$ids[1], db="all")$links$sra_bioproject_all)
getBPid

getSubmitter <- sapply(getBPid, function(x) extract_from_esummary(entrez_summary(db="bioproject", x[[1]][1]), "submitter_organization"))
getSubmitter

getSubmitterdf <- as.data.frame(getSubmitter)
getSubmitterdf$SRA <- rownames(getSubmitterdf)

# And now we add that info back to the relevant column, and replace the GEO calls. I honestly can't remember if I need to vectorise or not but whatever. 
allSRAFinal$finalSite <- NA

for (i in 1:nrow(getSubmitterdf)){
  allSRAFinal[allSRAFinal$SRA.Study %in% getSubmitterdf$SRA[i],]$finalSite <- getSubmitterdf$getSubmitter[i]    
}

allSRAFinal$finalSite <- coalesce(allSRAFinal$finalSite, allSRAFinal$Center.Name)

# Is GEO really gone???
length(unique(allSRAFinal$Center.Name))
length(unique(allSRAFinal$finalSite)) # Well that definitely increased, so that's good. 

siteTable <- allSRAFinal %>% count(finalSite, SRA.Study)

# I still need to work out what is in what country, but at least I have a single table... 
write.table(siteTable, file="final_Site_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")

# Anyhow, let's make sure we only do this once:
saveRDS(allSRAFinal, "allSRAFinalTissuesCenters.rds")
```

I decided that working out the country was easiest if done by hand, so:

```{r update-sampling-site-country}
allSRAFinal <- readRDS("allSRAFinalTissuesCenters.rds")
countryList <- read.csv("20240624_final_Site_descriptors.csv")

allSRAFinal$finalCountry <- NA

for (i in 1:nrow(countryList)){
  allSRAFinal[allSRAFinal$SRA.Study %in% countryList$SRA.Study[i],]$finalCountry <- countryList$Country[i]    
}

sort(table(allSRAFinal$finalCountry), decreasing=T)

```

And now my other nemesis... disease state. There's a lot of cancer info already in there, but what else is there?

```{r disease-status-1, include=T, eval=F}
# Start by checking the columns where disease status might hide, that we haven't looked at already:
names(allSRAFinal)
head(sort(table(allSRAFinal$disease), decreasing=T), n=20)
head(sort(table(allSRAFinal$disease_state), decreasing=T), n=20)
head(sort(table(allSRAFinal$Diagnosis), decreasing=T), n=20)
head(sort(table(allSRAFinal$clinical_information), decreasing=T), n=20)
head(sort(table(allSRAFinal$health_state), decreasing=T), n=20)
head(sort(table(allSRAFinal$DONOR_HEALTH_STATUS), decreasing=T), n=20)

# All of these make sense except for clinical_information, so... we go back to doing the same stuff we did with the population descriptors.
write.table(as.data.frame(table(allSRAFinal$disease)), file="disease_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")
write.table(as.data.frame(table(allSRAFinal$disease_state)), file="disease_state_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")
write.table(as.data.frame(table(allSRAFinal$Diagnosis)), file="Diagnosis_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")
write.table(as.data.frame(table(allSRAFinal$health_state)), file="health_state_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")
write.table(as.data.frame(table(allSRAFinal$DONOR_HEALTH_STATUS)), file="DONOR_HEALTH_STATUS_descriptors_raw.tsv", quote=F, row.names=F, sep="\t")
```

And now to check the intersections etc etc...

```{r disease-status-2}
# First, how independent are all of these:
table(is.na(allSRAFinal$disease))
table(is.na(allSRAFinal$Diagnosis))
table(is.na(allSRAFinal$disease_state))
table(is.na(allSRAFinal$health_state))
table(is.na(allSRAFinal$DONOR_HEALTH_STATUS))
      
# And now please tell me they don't overlap too much (F/F would be overlap):
table(is.na(allSRAFinal$disease), is.na(allSRAFinal$Diagnosis))
table(is.na(allSRAFinal$disease), is.na(allSRAFinal$disease_state))
table(is.na(allSRAFinal$disease), is.na(allSRAFinal$health_state)) # ARGH
table(is.na(allSRAFinal$disease), is.na(allSRAFinal$DONOR_HEALTH_STATUS)) #ARGH

table(is.na(allSRAFinal$disease_state), is.na(allSRAFinal$Diagnosis))
table(is.na(allSRAFinal$disease_state), is.na(allSRAFinal$health_state))
table(is.na(allSRAFinal$disease_state), is.na(allSRAFinal$DONOR_HEALTH_STATUS))

table(is.na(allSRAFinal$health_state), is.na(allSRAFinal$Diagnosis))
table(is.na(allSRAFinal$health_state), is.na(allSRAFinal$DONOR_HEALTH_STATUS))

table(is.na(allSRAFinal$Diagnosis), is.na(allSRAFinal$DONOR_HEALTH_STATUS))

# Checking the overlaps:
allSRAFinal[!is.na(allSRAFinal$disease) & !is.na(allSRAFinal$DONOR_HEALTH_STATUS),] # DONOR_HEALTH_STATUS wins, so when we coalesce it will go first.

allSRAFinal %>% count(disease, health_state) # Disease wins.

# Now we update terms:
diseaseTerms <- read.csv("20240624_disease_descriptors.csv")
allSRAFinal$finalDisease <- NA

# The laziest way to deal with the ordering here is to reorder the diseaseTerms file by the order we want:
# Diagnosis, health_state, disease_state, disease, DONOR_HEALTH_STATUS 
# That way as we move down it, we'll overwrite previously incorrect assignments:

diseaseTerms$coded.as <- factor(diseaseTerms$coded.as, levels=c("Diagnosis", "health_state", "disease_state", "disease", "DONOR_HEALTH_STATUS"))
diseaseTerms <- diseaseTerms[order(diseaseTerms$coded.as),]

for (i in 1:nrow(diseaseTerms)){
  allSRAFinal[allSRAFinal[,as.character(diseaseTerms$coded.as[i])] %in% diseaseTerms$Term[i],]$finalDisease <- diseaseTerms$grouping[i] 
}

# And we mustn't forget to add all the other cancer calls!
allSRAFinal[allSRAFinal$finalOrgan %in% "cancer",]$finalDisease <- "cancer"
table(allSRAFinal$finalDisease)

# Check that this actually worked...
allSRAFinal[!is.na(allSRAFinal$health_state) & !is.na(allSRAFinal$disease),]

# Yep looks good, now we can really analyse things:
saveRDS(allSRAFinal, "allSRAFinalTissuesCentersDisease.rds")

```
