---
title: "Merging the PCA with our calls, part 2 - quick exploration of data."
layout: "post"
tags: "diversity"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=F}
knitr::opts_knit$set(root.dir="~/OneDrive - St Vincent's Institute/Documents/RNA\ Diversity/", echo=T)
# But we also have to do this manually because R studio is stupid: 
setwd(knitr::opts_knit$get("root.dir"))
library(ggplot2)
library(data.table)
library(plyr)
library(reshape)
library(viridis)
library(tidyverse)
library(patchwork)
library(ggalluvial)
library(RColorBrewer)
library(ggrepel)
library(viridis)
library(thematic)
library(tidyverse)

options(device = "quartz")
```

Now that we have finished doing the programmatic grabbing from NCBI, let's do the analyses we want on the two datasets combined. The things we want to do are:

1. diversity in his data, after cleaning it down to biosample, 10 samples per study, and removing obvious cell lines (HeLa etc)
2. agreement between him and us, as either alluvial or confusion matrix - this can also lead to point 3 below.
3. sample swaps in Run IDs, but that's really just for my own personal edification - doesn't fit with the paper.

So let's start with 1, even though it's annoying. These are big files, so let's load them in as quickly as we need to and then throw them out... at least we know what columns we want to keep, this time around. We load in the rds with the country info from the previous script, and then we merge with Kasper, although we should not discard samples just because they have no info from him - we stick to all studies that passed our criteria, after we wipe the cell line column.

```{r read-in-original-data}
allSRAFilt <- readRDS("20250605_allSRAFilt_country.rds")

# So now we finally clean up to one row per biosample...
allSRAClean <- allSRAFilt %>% distinct(BioSample, .keep_all=TRUE)
dim(allSRAClean)

# A bit of tidying up to free up some memory etc...
rm(allSRA, allSRAFilt, allSRAcounts)
```

And now we read in Kasper's file, my SRA dump and merge, sanity checking the missingness in the data because otherwise it's all quite yikes:

```{r merge-with-kasper}
fromKasperBase <- read.csv("from_Kasper_suppTable3.csv", header=T)
bioSampleMappings <- read.csv("final_bioproject_biosample_runlist.csv", header=T)

# While R is geting its act together we need to do some manual concatenation in the shell:
# ( cat bioproject_biosample_runlist_to_66.csv ; tail -n +2 bioproject_biosample_runlist.csv ) >> bioproject_biosample_runlist_working.csv

dim(fromKasperBase)
head(fromKasperBase)
dim(bioSampleMappings)
head(bioSampleMappings)

# This can be an inner join now so we can sanity check... but we're clearly pulling a lot of IDs that are not included in recount, but are associated with the same project and are still transcriptomic somehow. It is above my pay grade to work out why. 
fromKasper <-inner_join(fromKasperBase, bioSampleMappings, join_by("sub_pop" == "Run")) 
dim(fromKasperBase) - dim(fromKasper) # only missing 1000, not bad!

# Anyhow, let's drop those annoying samples without pop label so we can drop them whenever we need to. 
fromKasper <- fromKasper %>%
  mutate(pred_pop = na_if(pred_pop, "")) %>%
  filter(!is.na(pred_pop))

# But this was at the run level, we gotta get down to the BioSample level...
length(unique(fromKasper$BioSample))
fromKasper <- fromKasper %>% distinct(BioSample, .keep_all=TRUE)
dim(fromKasper) #That's better

# Now we merge with allSRAClean to check some stuff...
dim(allSRAClean)
allSRAClean <- inner_join(allSRAClean, fromKasper, join_by("BioSample" == "BioSample"))
dim(allSRAClean)

nrow(allSRAClean) / nrow(fromKasper) # 92% of the Kasper rows were in our data, so that's a start

# Not bad... if we do the PCA and the tallying at this point, what does it look like? 
fromKasper %>% count(pred_pop)
allSRAClean %>% count(pred_pop)

allSRAClean %>% count(is.na(Cell_line))
popularCellLines <- allSRAClean %>% count(Cell_line, sort=T) 
dim(popularCellLines)

popularCellTypes <- allSRAClean %>% count(Cell_type, sort=T) 
dim(popularCellTypes)

# head(popularCellLines, n=200)
# head(popularCellTypes, n=200)

# Let's clean up, then - and also get rid of all the samples without a genotype prediction from their data
allSRAClean <- allSRAClean %>% 
  filter(is.na(Cell_line)) %>%
  filter(!is.na(pred_pop))
dim(allSRAClean) # This is the final number that match with some sort of population label.
allSRAClean %>% count(pred_pop) # And these are the pop numbers
```

From a quick look at this, it seems like most people do a decent job of filling in the cell type and cell line categories, so we'll go for the savage approach: drop everything in Kasper's file with a cell_line annotation and say that's as close as we can get to matching our far more detailed clean up. 

First we get ready to plot by defining some colour schemes, and then we need to do some mappings between them and us, so might as well read in allSRAFinal because that's what we need downstream anyhow and I forgot the colour levels. 

```{r prepare-to-plot}
allSRAFinal <- readRDS("20240901_allSRAFinal_for_plotting.rds")

scale_fill_geography <- function(...){
    ggplot2:::manual_scale('fill',  
        values = setNames(c('#9C8DC3', '#F3D78A', '#DB6968', '#60BC55', '#BCCC45', '#4D97CD', '#04C3C8', '#C69C3A', '#F8984E', '#8B96AD', '#FBC9C4'), levels(allSRAFinal$strictestGeography)), 
    )
}

scale_color_geography <- function(...){
    ggplot2:::manual_scale('color',  
        values = setNames(c('#9C8DC3', '#F3D78A', '#DB6968', '#60BC55', '#BCCC45', '#4D97CD', '#04C3C8', '#C69C3A', '#F8984E', '#8B96AD', '#FBC9C4'), levels(allSRAFinal$strictestGeography)), 
    )
}

# And now the fill scales for Kasper, which are basically ours, but smaller - we gotta reorder his factor levels so they match well:
allSRAClean$pred_pop <- factor(allSRAClean$pred_pop, levels = c("AFR", "EUR", "SAS", "EAS", "AMR", "")) 

# NA is grey50
scale_fill_kasper <- function(...){
    ggplot2:::manual_scale('fill',  
        values = setNames(c('#9C8DC3', '#DB6968', '#60BC55', '#4D97CD', '#F8984E', 'grey50'), levels(allSRAClean$pred_pop)), 
    )
}

scale_color_kasper <- function(...){
    ggplot2:::manual_scale('color',  
        values = setNames(c('#9C8DC3', '#DB6968', '#60BC55', '#4D97CD', '#F8984E', 'grey50'), levels(allSRAClean$pred_pop)), 
    )
}


raceColours <- brewer.pal(8,"Set2") #8 because we don't plot NAs

scale_color_race <- function(...){
    ggplot2:::manual_scale('color', 
                           values = setNames(raceColours, levels(allSRAFinal$strictestRace)), 
    )
}

scale_fill_race <- function(...){
    ggplot2:::manual_scale('fill', 
        values = setNames(raceColours, levels(allSRAFinal$strictestRace)), 
    )
}


theme_set(theme_bw(base_size = 6))
theme_update(axis.text=element_text(size=7))
theme_update(legend.key.size = unit(12, 'pt'), #change legend key size
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6),
        plot.title = element_text(size=8))
theme_update(plot.margin = unit(c(2, 2, 2, 2), "pt"))

# We also define a couple of plot layouts:
long3Design <- "
  1
  2
  3
"
```

Now we can finish poking around and do the actual plotting, and eventually merge this with other figure drafts. 

```{r plot-clean-data}
kasperProp <- allSRAClean %>% count(pred_pop) %>% mutate(freq = n/sum(n))

kasperPlot <- ggplot(kasperProp, aes(x = fct_inorder(pred_pop), y = n, fill = pred_pop)) +
  geom_bar(stat="identity") +
  ggtitle("Inferred genetic superpopulation\n(Razi et al, 2024)") +
  xlab(NULL) +
  ylab(NULL) +
  coord_flip() +
  scale_fill_kasper() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  theme(plot.title = element_text(hjust = 1)) 

kasperPlot
ggsave("kasper_filtered_allsra_by_region.pdf", height=3, width=6)
ggsave("kasper_filtered_allsra_by_region.png", height=3, width=6)

# We also plot the PCA 
kasperPCA <- ggplot(allSRAClean, aes(x = pc1, y = pc2, color=pred_pop )) +
  ggtitle(paste0("Genotype PCA (Razi et al, 2024)\n(n = ", nrow(allSRAClean), " samples)")) +
  geom_point(alpha = 0.5, size = 0.5) +
  ylab("PC2") +
  xlab("PC1") +
  theme(plot.title = element_text(hjust = 1)) +
  scale_color_kasper() +
  labs(color="Predicted\nSuperpopulation") +
  theme(legend.position = "none")
  # theme(legend.title.align = 1, legend.position = c(0.1,0.8), legend.background=element_blank()) +
  # theme(aspect.ratio=1)

kasperPCA

kasperPlot + kasperPCA +
  # plot_layout(design = long6Design, tag_level = 'new') +
    plot_annotation(tag_levels = 'A') 

ggsave("kasper_filtered_pca_by_region.pdf", height=3, width=6)
ggsave("kasper_filtered_pca_by_region.png", height=3, width=6)
```

And now we can start intersecting with our cleaned SRA dataset from the paper...

```{r merge-sra-annot-kasper}
allSRAFinalPC <- left_join(allSRAFinal, fromKasper, join_by("BioSample" == "BioSample"))

# Check the merges again:
dim(allSRAFinal)
dim(allSRAFinalPC)
dim(fromKasper)

# Let's check the quality of this merger... 
allSRAFinalPC %>% drop_na(BioProject) %>% nrow() # So much missing data... only 5k samples are in common in both datasets right now
allSRAFinalPC %>% drop_na(BioProject) %>% count(BioProject) %>% nrow()

# Are any of these incomplete?
mergedBPs <- allSRAFinalPC %>% drop_na(BioProject) %>% count(BioProject)
sraBPs <- allSRAFinalPC %>% count(SRA.Study)

bpChecks <- inner_join(mergedBPs, sraBPs, join_by("BioProject" == "SRA.Study"))
bpChecks %>% mutate(diff = n.y - n.x) # Mostly equal, but some weird missingness in some cases... Let's see how/if that changes when we get it to grab the complete datasets.

allSRAFinalPC %>% drop_na(BioProject) %>% count(SRA.Study == BioProject) # Only three mismatches, tho... 
allSRAFinalPC %>% drop_na(BioProject) %>% filter(SRA.Study != BioProject) # Terrifying. Checked them manually and they line up at the BioSample level and everything was deposited by Sanger (and they both have an PREJ Bioproject instead), but still, wtf?
                         
# How many of these are just pending?
recountBPs <- unique(fromKasperBase$pop)
processedBPs <- unique(bioSampleMappings$BioProject)
SRABPs <- unique(allSRAFinal$SRA.Study)

length(recountBPs)
length(processedBPs)
length(SRABPs)

table(SRABPs %in% recountBPs) # So we only have 94 studies in common with recount, not the full complement. Of course not. 
table(SRABPs %in% processedBPs) 
```

This could be better but that's what we have - how did they choose their samples? Anyhow, we'll start easy and ask, does having usable metadata change proportions of samples relative to the entire dataset?

```{r merge-sra-annot-kasper-2}
# First we explore the merge a little bit. This was a left join, so lots of pred_pop with NA, let's filter those out for tallying, because it's kind of unfair otherwise:
allSRAFinalPC %>% drop_na(BioProject) %>% count(strictestGeography)
allSRAFinalPC %>% drop_na(BioProject) %>% count(strictestRace)
allSRAFinalPC %>% drop_na(BioProject) %>% count(pred_pop) 

# Some confusion... that AMR category means what anyhow??? It's absolutely pointless at the genetics level, unless I've fucked something up above? 
allSRAFinalPC %>% drop_na(BioProject) %>% drop_na(strictestGeography) %>% count(strictestGeography, pred_pop)
allSRAFinalPC %>% drop_na(BioProject) %>% drop_na(strictestRace) %>% count(strictestRace, pred_pop)

# Anyhow let's answer the question
descriptorPredPop <- allSRAFinalPC %>% drop_na(BioProject) %>% count(pred_pop) %>% mutate(freq = n/sum(n)) %>% arrange()
allSRAPredPop <- allSRAClean %>% drop_na(BioProject) %>% count(pred_pop) %>% mutate(freq = n/sum(n)) %>% arrange(as.character(pred_pop))

geoPredPop <- allSRAFinalPC %>% drop_na(c(BioProject, strictestGeography)) %>% count(pred_pop) %>% mutate(freq = n/sum(n)) %>% arrange()

onion <- chisq.test(cbind(descriptorPredPop$n, allSRAPredPop$n)) # The difference is significant, but how significant, and in what direction?
onion
# For checking
onion$expected
onion$observed

geOnion <- chisq.test(cbind(geoPredPop$n, allSRAPredPop$n)) # The difference is significant, but how significant, and in what direction?
geOnion
# For checking
geOnion$expected
geOnion$observed

# Regardless of descriptor type
allPreds <- bind_rows(lst(descriptorPredPop, allSRAPredPop), .id='id') %>% 
  mutate(id = gsub("descriptorPredPop", "Has metadata\ndescriptor", id)) %>%
  mutate(id = gsub("allSRAPredPop", "All SRA samples", id))

descriptorFreqs <- ggplot(allPreds, aes(x=id, y=freq, fill=pred_pop)) +
  geom_bar(position="fill", stat="identity")+
    # theme_bw(13)+
    ggtitle("") +
    ylab('Proportion of samples')+
    xlab('')+
    scale_fill_kasper() +
    scale_y_continuous(breaks=c(0,0.5,1)) +
    guides(fill=guide_legend(title="Superpop label", label.position="left", nrow=2)) +
    theme(legend.title.align = 1, legend.direction = "horizontal", legend.background=element_blank(), legend.justification = c(0,0), legend.position="bottom", legend.title=element_blank())

descriptorFreqs
  ggsave("sra_kasper_merged.pdf", height=4, width=4)
  ggsave("sra_kasper_merged.png", height=4, width=4)

# What about only those samples with geographic descriptors?  
geoPreds <- bind_rows(lst(geoPredPop, allSRAPredPop), .id='id') %>% 
  mutate(id = gsub("geoPredPop", "Has geographic\ndescriptor", id)) %>%
  mutate(id = gsub("allSRAPredPop", "All SRA samples", id))

geoFreqs <- ggplot(geoPreds, aes(x=id, y=freq, fill=pred_pop)) +
  geom_bar(position="fill", stat="identity")+
    # theme_bw(13)+
    ggtitle("") +
    ylab('Proportion of samples')+
    xlab('')+
    scale_fill_kasper() +
    scale_y_continuous(breaks=c(0,0.5,1)) +
    guides(fill=guide_legend(title="Superpop label", label.position="left", nrow=2)) +
    theme(legend.title.align = 1, legend.direction = "horizontal", legend.background=element_blank(), legend.justification = c(0,0), legend.position="bottom", legend.title=element_blank())

geoFreqs
  ggsave("sra_kasper_merged_geo_only.pdf", height=4, width=4)
  ggsave("sra_kasper_merged_geo_only.png", height=4, width=4)  
  
  
# And now... an alluvial plot god give me strength
allSRAFinalPC %>% drop_na(c(strictestGeography, pred_pop)) %>% nrow()
allSRAFinalPC %>% drop_na(c(strictestRace, pred_pop)) %>% nrow()
allSRAFinalPC %>% drop_na(c(strictestGeography)) %>% nrow()
allSRAFinalPC %>% drop_na(c(strictestRace)) %>% nrow()
allSRAFinalPC %>% drop_na(c(pred_pop)) %>% nrow()

  
# The big empty box in the Superpop category is samples without a superpop
geoPCAAlluvial <- allSRAFinalPC %>% drop_na(c(strictestGeography)) %>% count(strictestGeography, pred_pop) %>%
  ggplot(data = .,
       aes(axis1 = strictestGeography, axis2 = pred_pop, y = n)) +
  scale_x_discrete(limits = c("Reported geographic descriptor", "Predicted superpop"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestGeography)) +
  scale_fill_geography() +
  geom_stratum(width=1/4) +
  ylab("Samples") +
  # geom_text(stat = "stratum", aes(label = after_stat(stratum)), size=2.5) +
  geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 2.5, direction = "y", nudge_x = .5) +
  theme_minimal(base_size = 6) +
  guides(fill=guide_legend(title="Descriptor", nrow=2)) +
  theme(legend.title = element_blank(), legend.position="bottom", legend.direction="horizontal")

geoPCAAlluvial

geoPCAAlluvialSlim <- allSRAFinalPC %>% drop_na(c(strictestGeography, pred_pop)) %>% count(strictestGeography, pred_pop) %>%
  ggplot(data = .,
       aes(axis2 = strictestGeography, axis1 = pred_pop, y = n)) +
  scale_x_discrete(limits = rev(c("Reported geographic descriptor", "Predicted superpop")), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = pred_pop)) +
  scale_fill_kasper() +
  geom_stratum(width=1/4) +
  ylab("Samples") +
  # geom_text(stat = "stratum", aes(label = after_stat(stratum)), size=2.5) +
  geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 2.5, direction = "y", nudge_x = .5) +
  theme_minimal(base_size = 6) +
  guides(fill=guide_legend(title="Descriptor", nrow=2)) +
  theme(legend.title = element_blank(), legend.position="bottom", legend.direction="horizontal")

geoPCAAlluvialSlim

# These are very hard to interpret and I'm just making them to sanity check what happens to the Hispanic category
racePCAAlluvial <- allSRAFinalPC %>% drop_na(c(strictestRace)) %>% count(strictestRace, pred_pop) %>%
  ggplot(data = .,
       aes(axis1 = strictestRace, axis2 = pred_pop, y = n)) +
  scale_x_discrete(limits = c("Reported racial descriptor", "Predicted superpop"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = strictestRace)) +
  scale_fill_race() +
  geom_stratum(width=1/4) +
  ylab("Samples") +
  # geom_text(stat = "stratum", aes(label = after_stat(stratum)), size=2.5) +
  geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 2.5, direction = "y", nudge_x = .5) +
  theme_minimal(base_size = 6) +
  guides(fill=guide_legend(title="Descriptor", nrow=2)) +
  theme(legend.title = element_blank(), legend.position="bottom", legend.direction="horizontal")

racePCAAlluvial

racePCAAlluvialSlim <- allSRAFinalPC %>% drop_na(c(strictestRace, pred_pop)) %>% count(strictestRace, pred_pop) %>%
  ggplot(data = .,
       aes(axis2 = strictestRace, axis1 = pred_pop, y = n)) +
  scale_x_discrete(limits = rev(c("Reported racial descriptor", "Predicted superpop")), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = pred_pop)) +
  scale_fill_kasper() +
  geom_stratum(width=1/4) +
  ylab("Samples") +
  # geom_text(stat = "stratum", aes(label = after_stat(stratum)), size=2.5) +
  geom_text_repel(stat = "stratum", aes(label = after_stat(stratum)), size = 2.5, direction = "y", nudge_x = .5) +
  theme_minimal(base_size = 6) +
  guides(fill=guide_legend(title="Descriptor", nrow=2)) +
  theme(legend.title = element_blank(), legend.position="bottom", legend.direction="horizontal")

racePCAAlluvialSlim

# Honestly the most interesting thing is the much greater proportion of NAs in these vs the geo ones. How did recount3 choose their samples???

geoFreqs + geoPCAAlluvial +
  plot_annotation(tag_levels = 'A') 
  ggsave("sra_kasper_all_alluvial.pdf", height=3, width=6)
  ggsave("sra_kasper_all_alluvial.png", height=3, width=6)

geoFreqs + geoPCAAlluvialSlim +
  plot_annotation(tag_levels = 'A') 
  ggsave("sra_kasper_slim_alluvial.pdf", height=3, width=6)
  ggsave("sra_kasper_slim_alluvial.png", height=3, width=6)  
```

And now we just make the confusion-like matrix and turn that into a heatmap. Amanda can choose what she likes best. 

```{r merge-sra-annot-kasper-3}
# Create matrix
confMatrix <- table(Predicted = allSRAFinalPC$strictestGeography, Actual = allSRAFinalPC$pred_pop)

# Convert to data frame and calculate percentages
df_cm <- melt(confMatrix)
df_cm$Percentage <- round(100 * df_cm$value / sum(df_cm$value), 1)

# Reverse the order of Predicted factor levels
df_cm$Predicted <- factor(df_cm$Predicted, levels = rev(unique(df_cm$Predicted)))

# Plot
confMatrixPlot <- ggplot(df_cm, aes(x = Actual, y = Predicted, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = ifelse(value == 0, "", paste0(value, "\n", Percentage, "%"))),
            size = 1.5, color = "black") +
  scale_fill_gradient(low = "white", high = "steelblue") +
  xlab("Inferred superpopulation\n(Razi et al, 2024)") + 
  ylab("Submitted descriptor\n(ancestral/geographic origin)") +
  theme(legend.position = "none")
  
confMatrixPlot

geoFreqs + geoPCAAlluvial + confMatrixPlot +
  plot_layout(design = long3Design, tag_level = 'new') +
  plot_annotation(tag_levels = 'A') 
  ggsave("sra_kasper_all_alluvial_conf_matrix.pdf", height=7, width=3)
  ggsave("sra_kasper_all_alluvial_conf_matrix.png", height=7, width=3)

geoFreqs + confMatrixPlot +
  # plot_layout(design = long3Design, tag_level = 'new') +
  plot_annotation(tag_levels = 'A') 
  ggsave("sra_kasper_all_conf_matrix.pdf", height=3, width=6)
  ggsave("sra_kasper_all_conf_matrix.png", height=3, width=6)

geoPCAAlluvial + confMatrixPlot +
  # plot_layout(design = long3Design, tag_level = 'new') +
  plot_annotation(tag_levels = 'A') 
  ggsave("sra_kasper_all_alluvial_conf_matrix.pdf", height=3, width=6)
  ggsave("sra_kasper_all_alluvial_conf_matrix.png", height=3, width=6)

geoPCAAlluvialSlim + confMatrixPlot +
  # plot_layout(design = long3Design, tag_level = 'new') +
  plot_annotation(tag_levels = 'A') 
  ggsave("sra_kasper_slim_alluvial_conf_matrix.pdf", height=3, width=6)
  ggsave("sra_kasper_slim_alluvial_conf_matrix.png", height=3, width=6)
```

I think that's all we need, really... no point in digging much deeper when it is all going to be mighty confusing anyhow. And since the actual plotting and integration happens in the next scripts, let's write out our merged outputs.

```{r write-output}
saveRDS(allSRAClean, file="20250508_allSRA_data_razi_merge.rds")
saveRDS(allSRAFinalPC, file="20250508_allSRAFinal_for_plotting_razi_merge.rds")
```
